---
title: "Full analysis"
subtitle: "Certificates at a glance, the headline findings, and the cotton case study"
---

```{r setup}
source("R/prep_dashboard.R")
```

::: {.callout-note appearance="minimal"}
**`r format(n_total, big.mark=",")` certificates** analysed, from the
[PPV&FR Authority register](https://plantauthority.gov.in/list-certificates).
Last data refresh: **`r last_updated`**. This page rebuilds every week.
:::

The Protection of Plant Varieties and Farmers' Rights Act (PPVFRA), 2001 set out to
protect breeders' rights while also recognising farmers as innovators. This page walks
through what the certification record shows. To ask these questions of a specific crop
yourself, use the [interactive explorer](index.qmd).

# Certificates at a glance

## Certificates issued over time

```{r}
f1 <- dd %>% count(issue_year)
year_2013_n <- first(f1$n[f1$issue_year == 2013], default = NA)
year_2022_n <- first(f1$n[f1$issue_year == 2022], default = NA)

plot_ly(f1, x = ~issue_year, y = ~n, type = "scatter", mode = "lines+markers",
        line = list(color = accent), marker = list(color = accent),
        hovertemplate = "%{x}: %{y} certificates<extra></extra>") %>%
  layout(title = "Total PPVFRA certificates issued, all crops",
         xaxis = list(title = "Year of certificate issue"),
         yaxis = list(title = "No. of certificates", tickformat = ",d"),
         annotations = list(
           list(x = 2013, y = year_2013_n, text = "2013", showarrow = TRUE, arrowhead = 2, ay = -35),
           list(x = 2022, y = year_2022_n, text = "2022", showarrow = TRUE, arrowhead = 2, ay = -35)))
```

Certificates spike in 2013 (`r comma(year_2013_n)` issued) and again in 2022
(`r comma(year_2022_n)`).

## Cotton, on its own

```{r}
f2 <- cot %>% count(issue_year)
plot_ly(f2, x = ~issue_year, y = ~n, type = "scatter", mode = "lines+markers",
        line = list(color = accent), marker = list(color = accent),
        hovertemplate = "%{x}: %{y} certificates<extra></extra>") %>%
  layout(title = "PPVFRA certificates issued, cotton",
         xaxis = list(title = "Year of certificate issue"),
         yaxis = list(title = "No. of certificates", tickformat = ",d"))
```

## What gets certified

```{r}
f3 <- d %>% count(crop_bucket) %>%
  mutate(crop_bucket = fct_relevel(fct_reorder(crop_bucket, n, .desc = TRUE), "Others", after = Inf))
ggplot(f3, aes(crop_bucket, n)) +
  geom_col(fill = accent) + geom_text(aes(label = comma(n)), vjust = -0.4, size = 3) +
  scale_y_continuous(labels = comma) +
  labs(title = "Certificates by crop group", x = NULL, y = "Certificates issued") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

Cereals lead by a wide margin. Vegetables are the second-largest group, followed by
fibre crops (which are almost all cotton).

## By variety category

```{r}
f4 <- d %>% filter(!is.na(variety_category)) %>% count(variety_category)
ggplot(f4, aes(variety_category, n, fill = variety_category)) +
  geom_col() + scale_fill_manual(values = cat_cols) + guides(fill = "none") +
  geom_text(aes(label = comma(n)), vjust = -0.4, size = 3) +
  scale_y_continuous(labels = comma) +
  labs(title = "Certificates by variety category", x = NULL, y = "Certificates issued")
```

Farmers' varieties are the largest single category overall. The sections below show that
this total sits very differently across applicant types and crops.

## Farmers' varieties over time

```{r}
f5 <- dd %>% filter(variety_category == "Farmer") %>% count(issue_year)
ggplot(f5, aes(issue_year, n)) +
  geom_col(fill = cat_cols[["Farmer"]]) + geom_text(aes(label = n), vjust = -0.4, size = 3) +
  scale_x_continuous(breaks = seq(yr_min, max(f5$issue_year), 1)) +
  scale_y_continuous(labels = comma) +
  labs(title = "Certificates issued for farmers' varieties annually",
       x = "Year of certificate issue", y = "Farmers' certificates") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Farmer variety certification has effectively ceased in recent years. The bars that once
built the largest single category on this page have trailed off to almost nothing.

## Public applicants, by category

```{r}
f6 <- dd %>% filter(applicant_category == "Public", !is.na(variety_category)) %>% count(issue_year, variety_category)
ggplot(f6, aes(issue_year, n, fill = variety_category)) + geom_col() +
  scale_fill_manual(values = cat_cols) +
  scale_x_continuous(breaks = seq(yr_min, max(f6$issue_year), 2)) +
  scale_y_continuous(labels = comma) +
  labs(title = "Certificates to public applicants, by category",
       x = "Year of certificate issue", y = "Certificates", fill = "Category")
```

## Private applicants, by category

```{r}
f7 <- dd %>% filter(applicant_category == "Private", !is.na(variety_category)) %>% count(issue_year, variety_category)
ggplot(f7, aes(issue_year, n, fill = variety_category)) + geom_col() +
  scale_fill_manual(values = cat_cols) +
  scale_x_continuous(breaks = seq(yr_min, max(f7$issue_year), 2)) +
  scale_y_continuous(labels = comma) +
  labs(title = "Certificates to private applicants, by category",
       x = "Year of certificate issue", y = "Certificates", fill = "Category")
```

## Ownership concentration beyond cotton

```{r}
top_firms_bar(d, title = "Major private applicants, all crops")
```

Market concentration is high across all crops, not just cotton. A small number of
companies dominate private variety certification in India. Cotton, covered next, is the
most extreme version of this pattern.

# The headline findings

::: {.callout-note appearance="simple"}
These four charts carry the argument of this page. A reader who only looks at this section
should leave knowing the story.
:::

## Applicant category, cotton vs all crops

```{r}
f8 <- bind_rows(d %>% count(sector) %>% mutate(set = "All crops"),
                cot %>% count(sector) %>% mutate(set = "Cotton")) %>% filter(!is.na(sector))
ggplot(f8, aes(sector, n, fill = set)) +
  geom_col(position = "dodge") + scale_fill_manual(values = set_cols) +
  geom_text(aes(label = comma(n)), position = position_dodge(width = 0.9), vjust = -0.4, size = 3) +
  scale_y_continuous(labels = comma) +
  labs(title = "Applicant category: cotton vs all crops",
       x = "Applicant category", y = "Certificates", fill = NULL)
```

This is the central finding. Farmers hold the most certificates of any applicant type
across all crops combined. In cotton, that order inverts almost completely: private firms
hold nearly every certificate, and farmers are left with a handful. The applicant category
that dominates the register overall is nearly absent in India's most commercial crop.

## Cumulative ownership over time, all crops

```{r}
f_cum <- dd %>% filter(!is.na(sector)) %>% count(issue_year, sector) %>%
  arrange(issue_year) %>% group_by(sector) %>% mutate(cum = cumsum(n)) %>% ungroup()
ggplot(f_cum, aes(issue_year, cum, color = sector)) +
  geom_line(linewidth = 1.1) + geom_point(size = 1.5) +
  scale_color_manual(values = sector_cols) +
  scale_x_continuous(breaks = seq(yr_min, max(f_cum$issue_year), 2)) +
  scale_y_continuous(labels = comma) +
  labs(title = "Cumulative certificates by applicant category, all crops",
       x = "Year of certificate issue", y = "Cumulative certificates", color = NULL)
```

This is the inversion in the previous chart, unfolding year by year rather than as a single
endpoint. Watch where the lines separate: that is the moment the structural divergence
between applicant types took hold.

## Cotton applicant composition over time

```{r}
f10 <- cot %>% filter(!is.na(sector)) %>% count(issue_year, sector)
plot_ly(f10, x = ~issue_year, y = ~n, color = ~sector, colors = sector_cols,
        type = "scatter", mode = "lines+markers") %>%
  layout(title = "Cotton applicant composition over time",
         xaxis = list(title = "Year of certificate issue"),
         yaxis = list(title = "No. of certificates", tickformat = ",d"))
```

Private applicants do not just dominate cotton certification, they have driven almost
every year of its growth. Public, SAU, and farmer lines stay close to flat by comparison.

## Who are the private cotton firms?

```{r}
top_firms_bar(cot, title = "Major private applicants, cotton")
```

A small number of firms, led by Nuziveedu, hold most of the private cotton certificates.
Varietal control in India's most commercial crop rests with a few companies.

# Cotton in detail

The headline findings established that cotton inverts the usual applicant pattern. This
section supports that finding with more detail on how.

## Ploidy by applicant

```{r}
f9 <- cot %>% filter(!is.na(ploidy), !is.na(sector)) %>% count(sector, ploidy)
ggplot(f9, aes(sector, n, fill = ploidy)) +
  geom_col(position = "dodge") + scale_fill_manual(values = ploidy_cols) +
  geom_text(aes(label = n), position = position_dodge(width = 0.9), vjust = -0.4, size = 3) +
  scale_y_continuous(labels = comma) +
  labs(title = "Cotton certificates by ploidy", x = "Applicant category",
       y = "No. of PVP certificates", fill = NULL)
```

Private cotton certification is mostly tetraploid, the hybrid and Bt lines. Diploid desi
cotton makes up a small share for every type of applicant.

## Variety category by applicant

```{r}
f11 <- cot %>% filter(!is.na(variety_category), !is.na(sector)) %>% count(variety_category, sector)
ggplot(f11, aes(variety_category, n, fill = sector)) + geom_col() +
  scale_fill_manual(values = sector_cols) +
  geom_text(aes(label = n), position = position_stack(vjust = 0.5), size = 3) +
  scale_y_continuous(labels = comma) +
  labs(title = "Cotton certificates by variety category and applicant",
       x = "Variety category", y = "Certificates", fill = NULL) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

## Private cotton, by category over time

```{r}
f12 <- cot %>% filter(applicant_category == "Private", !is.na(variety_category)) %>% count(issue_year, variety_category)
ggplot(f12, aes(issue_year, n, fill = variety_category)) + geom_col() +
  scale_fill_manual(values = cat_cols) +
  geom_text(aes(label = n), position = position_stack(vjust = 0.5), size = 3) +
  scale_x_continuous(breaks = seq(yr_min, max(f12$issue_year), 2)) +
  scale_y_continuous(labels = comma) +
  labs(title = "Private cotton certificates by category",
       x = "Year of certificate issue", y = "Certificates", fill = "Category")
```

---

::: {.callout-tip appearance="minimal"}
Full pipeline (scrape, clean, figures, and this page) is in the
[project repository](https://github.com/ambika-subash/pvp-certification-explorer).
:::
