---
title: "Full analysis"
subtitle: "Certificates at a glance, the headline findings, and the cotton case study"
---

```{r setup}
source("R/prep_dashboard.R")

# ---- figures quoted inline in the prose below, computed once here so a
# number mentioned in one caption can never drift out of sync with another
# caption, a chart, or the weekly data refresh ----
register_start_year <- min(d$issue_year, na.rm = TRUE)
register_end_year   <- max(d$issue_year, na.rm = TRUE)
register_years_span <- register_end_year - register_start_year + 1

f1 <- dd %>% count(issue_year)
overall_peak_year <- f1$issue_year[which.max(f1$n)]

f2 <- cot %>% count(issue_year)
cotton_peak_year <- f2$issue_year[which.max(f2$n)]

edv_all         <- d %>% filter(variety_category == "EDV")
edv_total       <- nrow(edv_all)
edv_crops       <- paste(sort(unique(as.character(edv_all$crop_bucket))), collapse = ", ")
edv_years       <- sort(unique(edv_all$issue_year))
edv_yr_txt      <- if (length(edv_years) <= 1) paste(edv_years, collapse = "") else paste(range(edv_years), collapse = "-")
edv_all_cotton  <- edv_total > 0 && all(edv_all$is_cotton, na.rm = TRUE)
edv_all_private <- edv_total > 0 && all(edv_all$applicant_category == "Private", na.rm = TRUE)

cotton_total     <- nrow(cot)
cotton_private_n <- sum(cot$applicant_category == "Private", na.rm = TRUE)
cotton_farmer_n  <- sum(cot$sector == "Farmer", na.rm = TRUE)

cotton_vck         <- cot %>% filter(variety_category == "Extant (VCK)")
cotton_vck_total   <- nrow(cotton_vck)
cotton_vck_private <- sum(cotton_vck$applicant_category == "Private", na.rm = TRUE)
cotton_vck_public  <- sum(cotton_vck$applicant_category == "Public", na.rm = TRUE)

cotton_firms_ranked <- top_firms_ranked(cot)
top5_pct_sum   <- round(sum(cotton_firms_ranked$pct[seq_len(min(5, nrow(cotton_firms_ranked)))]), 1)
top5_firms_txt <- top_firms_prose(cotton_firms_ranked, 5)

sau_cotton_years <- cot %>% filter(sector == "SAU") %>% pull(issue_year)
sau_cotton_last  <- if (length(sau_cotton_years)) max(sau_cotton_years, na.rm = TRUE) else NA_integer_
```

[![DOI](https://zenodo.org/badge/1112427994.svg)](https://doi.org/10.5281/zenodo.21140353)

::: {.callout-note appearance="minimal"}
**`r format(n_total, big.mark=",")` certificates** analysed, from the
[PPV&FR Authority register](https://plantauthority.gov.in/list-certificates).
Last data refresh: **`r last_updated`**. This page rebuilds every week.
:::

The Protection of Plant Varieties and Farmers' Rights Act (PPVFRA), 2001 was designed as a
sui generis intellectual property regime recognising three classes of right-holders:
breeders, researchers, and farmers. The certification record covers every plant variety
protection certificate issued since `r register_start_year`. This page asks who has actually benefited from
the Act's operation. Cotton receives separate treatment throughout: it is India's only
commercially cultivated genetically modified crop, and the site where private
concentration is most complete.

To ask these questions of a specific crop yourself, use the
[interactive explorer](index.qmd).

::: {.callout-note appearance="minimal" icon="false"}
**About this data.** A certificate under the PPVFRA confers intellectual property rights
over a registered plant variety: the right to control commercial sale, reproduction, and
licensing for a fixed term. This register records certificates granted, not applications
filed or rejected. Certificates are issued to five applicant categories: Private firms,
State Agricultural Universities (SAUs), Public institutions, Farmers, and Individual
Breeders.
:::

::: {.callout-note collapse="true" appearance="simple"}
## Methodology and data notes

**Source.** The PPV&FR Authority's public certificate register:
[plantauthority.gov.in/list-certificates](https://plantauthority.gov.in/list-certificates).

**Update frequency.** Weekly, every Sunday, via an automated pipeline that re-scrapes the
full register, rebuilds the dataset, and republishes this site.

**Coverage.** Every certificate granted since `r register_start_year`, through the most
recent successful weekly refresh.

**What's excluded.** Applications that were withdrawn or rejected are not in this
register and are not represented here. This page only covers certificates that were
actually granted.

**Known limitations.** Company names are merged across spelling variants using a
name-matching heuristic (`canon_company()` / `alias_company()`), not a corporate-registry
lookup, so a genuine typo or an unlisted acronym could still cause one real firm to
appear as two. Every figure quoted in the prose below (firm percentages, EDV and VCK
counts, cotton applicant splits) is computed live from the current dataset at render
time, not hardcoded, so it updates automatically with each weekly refresh. Full
methodology is documented in the
[project repository](https://github.com/ambika-subash/pvp-certification-explorer#readme).
:::

# Certificates at a glance

## Certificates issued over time

```{r}
year_2013_n <- first(f1$n[f1$issue_year == 2013], default = NA)
year_2022_n <- first(f1$n[f1$issue_year == 2022], default = NA)

plot_ly(f1, x = ~issue_year, y = ~n, type = "scatter", mode = "lines+markers",
        line = list(color = accent), marker = list(color = accent),
        hovertemplate = "%{x}: %{y} certificates<extra></extra>") %>%
  layout(title = "Total PPVFRA certificates issued, all crops",
         xaxis = list(title = "Year of certificate issue"),
         yaxis = list(title = "No. of certificates", tickformat = ",d"),
         annotations = list(
           list(x = 2013, y = year_2013_n, text = "2013", showarrow = TRUE, arrowhead = 2, ay = -35),
           list(x = 2022, y = year_2022_n, text = "2022", showarrow = TRUE, arrowhead = 2, ay = -35)))
```

Certification grew slowly in the Act's early years, peaked in 2013-14 due to a surge in
NGO-facilitated farmer variety registrations, then fell during the Covid years.
`r if (overall_peak_year == register_end_year) paste0("The most recent year on record, ", register_end_year, ", is the highest annual total yet.") else paste0("Certification has partly recovered since, though ", register_end_year, " has not yet matched the ", overall_peak_year, " peak.")`
The recent recovery is dominated by private applicants, particularly in cotton.

## Cotton, on its own

```{r}
plot_ly(f2, x = ~issue_year, y = ~n, type = "scatter", mode = "lines+markers",
        line = list(color = accent), marker = list(color = accent),
        hovertemplate = "%{x}: %{y} certificates<extra></extra>") %>%
  layout(title = "PPVFRA certificates issued, cotton",
         xaxis = list(title = "Year of certificate issue"),
         yaxis = list(title = "No. of certificates", tickformat = ",d"))
```

Cotton variety certification has accelerated sharply in recent years.
`r if (cotton_peak_year == max(cot$issue_year, na.rm = TRUE)) paste0("The most recent year on record, ", cotton_peak_year, ", is the highest annual total yet.") else paste0(cotton_peak_year, " remains the highest annual total on record.")`
Private firms account for virtually all of this growth.
`r if (!is.na(sau_cotton_last)) paste0("SAU registrations, last active in ", sau_cotton_last, ", have not recovered.") else "SAU registrations have not recovered."`

## What gets certified

```{r}
f3 <- d %>% count(crop_bucket) %>%
  mutate(crop_bucket = fct_relevel(fct_reorder(crop_bucket, n, .desc = TRUE), "Others", after = Inf))
ggplot(f3, aes(crop_bucket, n)) +
  geom_col(fill = accent) + geom_text(aes(label = comma(n)), vjust = -0.4, size = 3) +
  scale_y_continuous(labels = comma) +
  labs(title = "Certificates by crop group", x = NULL, y = "Certificates issued") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

Cereals dominate, driven by rice registrations. Vegetables are the second-largest group,
ahead of fibre crops (which are entirely cotton). Cotton's outsized influence on the pages
that follow reflects its commercial significance and the intensity of private breeding
activity concentrated in a single crop, not its raw certificate count.

## By variety category

```{r}
f4 <- d %>% filter(!is.na(variety_category)) %>% count(variety_category)
ggplot(f4, aes(variety_category, n, fill = variety_category)) +
  geom_col() + scale_fill_manual(values = cat_cols) + guides(fill = "none") +
  geom_text(aes(label = comma(n)), vjust = -0.4, size = 3) +
  scale_y_continuous(labels = comma) +
  labs(title = "Certificates by variety category", x = NULL, y = "Certificates issued")
```

Farmers' varieties are the largest single category in aggregate, but this number requires
qualification. Farmer registrations are concentrated in traditional food crops, confer no
right to sell branded seed, and generate no royalties. In commercially significant crops,
particularly cotton, the distribution inverts. Essentially Derived Varieties (EDVs), the
category designed to trigger benefit-sharing when registered varieties are used as the
basis for derivative breeding, account for just `r edv_total` certificates across the
entire dataset.
`r if (edv_all_cotton) paste0("All ", edv_total, " are in cotton.") else ""`
`r if (edv_all_private) paste0("All ", edv_total, " are privately held.") else ""`

## Farmers' varieties over time

```{r}
f5 <- dd %>% filter(variety_category == "Farmer") %>% count(issue_year)
ggplot(f5, aes(issue_year, n)) +
  geom_col(fill = cat_cols[["Farmer"]]) + geom_text(aes(label = n), vjust = -0.4, size = 3) +
  scale_x_continuous(breaks = seq(yr_min, max(f5$issue_year), 1)) +
  scale_y_continuous(labels = comma) +
  labs(title = "Certificates issued for farmers' varieties annually",
       x = "Year of certificate issue", y = "Farmers' certificates") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

farmer_peak_year        <- f5$issue_year[which.max(f5$n)]
farmer_last_active_year <- max(f5$issue_year[f5$n > 0], na.rm = TRUE)
farmer_new_variety_n    <- dd %>% filter(sector == "Farmer", variety_category == "New") %>% nrow()
```

Farmer variety certification peaked in `r farmer_peak_year`, driven by NGO-facilitated
bulk registrations of traditional landraces. Since then it has declined sharply, with the
most recent issuance in `r farmer_last_active_year`.
`r if (farmer_new_variety_n == 0) "No farmer has received a certificate for a new variety under the Act." else paste0(farmer_new_variety_n, " farmer(s) have received a certificate for a new variety under the Act.")`

## Public applicants, by category

```{r}
f6 <- dd %>% filter(applicant_category == "Public", !is.na(variety_category)) %>% count(issue_year, variety_category)
ggplot(f6, aes(issue_year, n, fill = variety_category)) + geom_col() +
  scale_fill_manual(values = cat_cols) +
  scale_x_continuous(breaks = seq(yr_min, max(f6$issue_year), 2)) +
  scale_y_continuous(labels = comma) +
  labs(title = "Certificates to public applicants, by category",
       x = "Year of certificate issue", y = "Certificates", fill = "Category")
```

Public institution registrations are concentrated in extant and notified variety
categories, reflecting existing varieties registered for protection rather than newly
developed commercial innovations. ICAR's large registration count reflects Green
Revolution-era breeding momentum rather than contemporary activity. New variety
registrations by public institutions have declined significantly since the mid-2010s.

## Private applicants, by category

```{r}
f7 <- dd %>% filter(applicant_category == "Private", !is.na(variety_category)) %>% count(issue_year, variety_category)
ggplot(f7, aes(issue_year, n, fill = variety_category)) + geom_col() +
  scale_fill_manual(values = cat_cols) +
  scale_x_continuous(breaks = seq(yr_min, max(f7$issue_year), 2)) +
  scale_y_continuous(labels = comma) +
  labs(title = "Certificates to private applicants, by category",
       x = "Year of certificate issue", y = "Certificates", fill = "Category")
```

Private applicants register predominantly new varieties and Varieties of Common Knowledge
(VCKs). The VCK category is significant: these are varieties already in the public domain
that are converted into private assets through registration. Of `r cotton_vck_total` cotton
VCKs in the dataset, `r cotton_vck_private` are privately held and `r cotton_vck_public`
`r ifelse(cotton_vck_public == 1, "is", "are")` held by a public institution.

## Ownership concentration beyond cotton

```{r}
top_firms_bar(d, title = "Major private applicants, all crops")
```

Market concentration is high across all crops, not just cotton. A small number of
companies dominate private variety certification in India. Cotton, covered next, is the
most extreme version of this pattern.

# The headline findings

::: {.callout-note appearance="simple"}
These charts carry the argument of this page. A reader who only looks at this section
should leave knowing the story.
:::

## Applicant category, cotton vs all crops

```{r}
f8 <- bind_rows(d %>% count(sector) %>% mutate(set = "All crops"),
                cot %>% count(sector) %>% mutate(set = "Cotton")) %>% filter(!is.na(sector))
ggplot(f8, aes(sector, n, fill = set)) +
  geom_col(position = "dodge") + scale_fill_manual(values = set_cols) +
  geom_text(aes(label = comma(n)), position = position_dodge(width = 0.9), vjust = -0.4, size = 3) +
  scale_y_continuous(labels = comma) +
  labs(title = "Applicant category: cotton vs all crops",
       x = "Applicant category", y = "Certificates", fill = NULL)
```

Across all crops, farmers hold the largest share of certificates, driven by
NGO-facilitated registrations of traditional varieties in food crops. In cotton the
distribution inverts. Private firms hold `r comma(cotton_private_n)` of
`r comma(cotton_total)` cotton certificates. Farmers hold `r cotton_farmer_n`.
Transgenic hybrid production requires trait licensing agreements, backcross breeding
infrastructure, and biosafety containment facilities. These are not administrative
barriers reducible to policy adjustment; they are the technical conditions of production
in commercial cotton, and they are inaccessible to farmers as a class.

## Cumulative ownership over time, all crops

```{r}
f_cum <- dd %>% filter(!is.na(sector)) %>% count(issue_year, sector) %>%
  arrange(issue_year) %>% group_by(sector) %>% mutate(cum = cumsum(n)) %>% ungroup()
ggplot(f_cum, aes(issue_year, cum, color = sector)) +
  geom_line(linewidth = 1.1) + geom_point(size = 1.5) +
  scale_color_manual(values = sector_cols) +
  scale_x_continuous(breaks = seq(yr_min, max(f_cum$issue_year), 2)) +
  scale_y_continuous(labels = comma) +
  labs(title = "Cumulative certificates by applicant category, all crops",
       x = "Year of certificate issue", y = "Cumulative certificates", color = NULL)
```

This is the inversion in the previous chart, unfolding year by year rather than as a single
endpoint. Private firms have dominated cumulative certifications throughout the register's
history. The gap between private and all other categories has widened consistently since
2014, driven largely by a surge in cotton registrations. Watch where the lines separate:
that is the moment the structural divergence between applicant types took hold.

## Cotton applicant composition over time

```{r}
f10 <- cot %>% filter(!is.na(sector)) %>% count(issue_year, sector)
plot_ly(f10, x = ~issue_year, y = ~n, color = ~sector, colors = sector_cols,
        type = "scatter", mode = "lines+markers") %>%
  layout(title = "Cotton applicant composition over time",
         xaxis = list(title = "Year of certificate issue"),
         yaxis = list(title = "No. of certificates", tickformat = ",d"))
```

Private applicants do not just dominate cotton certification, they have driven almost
every year of its growth. Public, SAU, and farmer lines stay close to flat by comparison.

## Who are the private cotton firms?

```{r}
top_firms_bar(cot, title = "Major private applicants, cotton")
```

In cotton, five firms account for `r top5_pct_sum` percent of all private certificates
issued between `r min(cot$issue_year, na.rm = TRUE)` and `r max(cot$issue_year, na.rm = TRUE)`:
`r top5_firms_txt`. These figures cover direct variety registrations only. Trait-level control through
sublicensing, particularly Monsanto/Bayer's Bt and BGII events embedded across virtually
all tetraploid registrations, is not visible in registration data and is not captured
here.

## How concentrated is private cotton ownership?

```{r}
priv_cot_firms <- cot %>% filter(applicant_category == "Private") %>%
  mutate(key = alias_company(canon_company(applicant))) %>%
  count(key, name = "n") %>% arrange(n)
lorenz_df <- priv_cot_firms %>%
  mutate(cum_firms = row_number() / n(), cum_certs = cumsum(n) / sum(n)) %>%
  select(cum_firms, cum_certs) %>%
  bind_rows(tibble(cum_firms = 0, cum_certs = 0), .)
gini_cotton <- gini(priv_cot_firms$n)

ggplot(lorenz_df, aes(cum_firms, cum_certs)) +
  geom_area(fill = sector_cols[["Private"]], alpha = 0.15) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = muted) +
  geom_line(color = sector_cols[["Private"]], linewidth = 1.1) +
  scale_x_continuous(labels = percent, limits = c(0, 1)) +
  scale_y_continuous(labels = percent, limits = c(0, 1)) +
  labs(title = "Concentration of private cotton certificates among firms",
       x = "Share of firms, smallest to largest", y = "Cumulative share of certificates")
```

The dashed line is perfect equality, where every firm holds the same share. The further
the solid line bows away from it, the more concentrated ownership is. The Gini
coefficient here is **`r round(gini_cotton, 2)`**, on a scale from 0 (every firm holds an
equal share) to 1 (one firm holds everything).

## Public vs private innovation in cotton

```{r}
f_new_pubpriv <- cot %>% filter(variety_category == "New", sector %in% c("Public", "Private")) %>%
  count(issue_year, sector)
ggplot(f_new_pubpriv, aes(issue_year, n, color = sector)) +
  geom_line(linewidth = 1.1) + geom_point(size = 1.5) +
  scale_color_manual(values = sector_cols) +
  scale_x_continuous(breaks = seq(yr_min, max(f_new_pubpriv$issue_year, na.rm = TRUE), 2)) +
  scale_y_continuous(labels = comma) +
  labs(title = "New cotton variety registrations, public vs private",
       x = "Year of certificate issue", y = "New variety certificates", color = NULL)
```

This excludes VCKs and extant varieties, leaving only registrations of genuinely new
material. It isolates active breeding from the registration of already-existing
varieties, which is a separate process covered next.

# Cotton in detail

The headline findings established that cotton inverts the usual applicant pattern. This
section supports that finding with more detail on how.

## Ploidy by applicant

```{r}
f9 <- cot %>% filter(!is.na(ploidy), !is.na(sector)) %>% count(sector, ploidy)
ggplot(f9, aes(sector, n, fill = ploidy)) +
  geom_col(position = "dodge") + scale_fill_manual(values = ploidy_cols) +
  geom_text(aes(label = n), position = position_dodge(width = 0.9), vjust = -0.4, size = 3) +
  scale_y_continuous(labels = comma) +
  labs(title = "Cotton certificates by ploidy", x = "Applicant category",
       y = "No. of PVP certificates", fill = NULL)
```

Private cotton certification is almost entirely tetraploid, the hybrid and Bt lines
preferred by the textile industry for their longer staple length. Diploid or desi cotton,
historically adapted to low-input smallholder conditions, has been abandoned by both
private and public breeders. SAUs show a more balanced ploidy profile, reflecting a
residual public mandate to maintain crop diversity, but SAU cotton registrations
`r if (!is.na(sau_cotton_last) && sau_cotton_last < max(cot$issue_year, na.rm = TRUE)) paste0("have not recurred since ", sau_cotton_last, ".") else "continue only sporadically."`

```{r}
farmer_cot <- cot %>% filter(sector == "Farmer") %>%
  select(issue_year, variety_category, crop, applicant)
```

::: {.callout-note appearance="simple"}
**The farmer presence in cotton, in full.** `r nrow(farmer_cot)` certificate(s) out of
`r comma(nrow(cot))` cotton certificates were issued to a farmer applicant.
`r if (nrow(farmer_cot) >= 1) paste0("The first was issued in ", farmer_cot$issue_year[1], ", category ", as.character(farmer_cot$variety_category[1]), ", crop ", farmer_cot$crop[1], ".") else ""`
:::

## Variety category by applicant

```{r}
f11 <- cot %>% filter(!is.na(variety_category), !is.na(sector)) %>% count(variety_category, sector)
ggplot(f11, aes(variety_category, n, fill = sector)) + geom_col() +
  scale_fill_manual(values = sector_cols) +
  geom_text(aes(label = n), position = position_stack(vjust = 0.5), size = 3) +
  scale_y_continuous(labels = comma) +
  labs(title = "Cotton certificates by variety category and applicant",
       x = "Variety category", y = "Certificates", fill = NULL) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

Private firms register primarily new varieties and VCKs. Public institutions register
mostly extant and notified varieties. Farmers are confined to the farmers' variety
category. EDVs appear only in the private column. Only `r edv_total` EDV certificates have
been issued nationally across `r register_years_span` years of the Act's operation,
`r if (edv_all_cotton) "all in cotton" else "spread across crops"`. Whether firms
avoid the EDV category because of its benefit-sharing obligations, or register derivative
varieties as new varieties by meeting the DUS distinctness threshold through minor
morphological modification, the registration data cannot resolve.

## Every EDV certificate ever issued

```{r}
f_edv <- edv_all %>% count(issue_year)
edv_axis_max <- max(d$issue_year, na.rm = TRUE)
ggplot(f_edv, aes(issue_year, n)) +
  geom_segment(aes(xend = issue_year, y = 0, yend = n), color = cat_cols[["EDV"]], linewidth = 1) +
  geom_point(size = 4, color = cat_cols[["EDV"]]) +
  geom_text(aes(label = n), vjust = -1.2, size = 3.5) +
  scale_x_continuous(breaks = seq(yr_min, edv_axis_max, 1),
                      limits = c(yr_min - 1, edv_axis_max + 1)) +
  scale_y_continuous(limits = c(0, max(f_edv$n) * 1.3), breaks = pretty_breaks()) +
  labs(title = "Essentially Derived Variety (EDV) certificates, all years, all crops",
       x = "Year of certificate issue", y = "EDV certificates issued") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

`r edv_total` EDV certificates have been issued nationally, all in `r edv_crops`,
concentrated in `r edv_yr_txt`. No EDV has been registered before or after that window, in
any crop.

## Testing the VCK enclosure sequence

```{r}
f_vck_new <- cot %>% filter(applicant_category == "Private", variety_category %in% c("New", "Extant (VCK)")) %>%
  count(issue_year, variety_category)
ggplot(f_vck_new, aes(issue_year, n, fill = variety_category)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = cat_cols) +
  scale_x_continuous(breaks = seq(yr_min, max(f_vck_new$issue_year, na.rm = TRUE), 2)) +
  scale_y_continuous(labels = comma) +
  labs(title = "Private cotton: VCK vs New variety registrations over time",
       x = "Year of certificate issue", y = "Certificates", fill = "Category")
```

One reading of the VCK category is a sequential strategy: register existing
common-knowledge varieties first, then shift to new variety development once that
low-cost registration opportunity is exhausted.

## Private cotton, by category over time

```{r}
f12 <- cot %>% filter(applicant_category == "Private", !is.na(variety_category)) %>% count(issue_year, variety_category)
ggplot(f12, aes(issue_year, n, fill = variety_category)) + geom_col() +
  scale_fill_manual(values = cat_cols) +
  geom_text(aes(label = n), position = position_stack(vjust = 0.5), size = 3) +
  scale_x_continuous(breaks = seq(yr_min, max(f12$issue_year), 2)) +
  scale_y_continuous(labels = comma) +
  labs(title = "Private cotton certificates by category",
       x = "Year of certificate issue", y = "Certificates", fill = "Category")

private_cotton_totals   <- f12 %>% group_by(issue_year) %>% summarise(n = sum(n), .groups = "drop")
private_cotton_peak_yr  <- private_cotton_totals$issue_year[which.max(private_cotton_totals$n)]
peak_new_share <- f12 %>% filter(issue_year == private_cotton_peak_yr) %>%
  mutate(share = n / sum(n)) %>% filter(variety_category == "New") %>% pull(share)
peak_new_share <- if (length(peak_new_share)) peak_new_share else 0
```

Private cotton certification shifted over time from VCK registrations toward new variety
registrations.
`r if (peak_new_share > 0.5) paste0("In ", private_cotton_peak_yr, ", the busiest year on record, most certificates were New variety registrations rather than VCKs.") else paste0("In ", private_cotton_peak_yr, ", the busiest year on record, VCK and New registrations were close to evenly split.")`
`r if (edv_total > 0) paste0("EDV registrations appeared briefly in ", edv_yr_txt, " and have not recurred.") else "No EDV registrations have occurred among private cotton applicants."`

---

::: {.callout-tip appearance="minimal"}
Data source: Protection of Plant Varieties and Farmers' Rights Authority public
certificate register, scraped weekly. Applicant categories are normalised from raw
register entries. All figures use year of certificate issue. Full pipeline is open source
and available in the
[project repository](https://github.com/ambika-subash/pvp-certification-explorer).
Built with R and Quarto.
:::

::: {style="margin-top:1.5rem;padding:1.25rem 1.5rem;background:#f9fafb;border:1px solid #e5e7eb;border-radius:8px;font-size:0.82rem;color:#4b5563;line-height:1.7"}
**License.** Code: [MIT License](https://github.com/ambika-subash/pvp-certification-explorer/blob/main/LICENSE).
Data and content: [CC BY 4.0](https://github.com/ambika-subash/pvp-certification-explorer/blob/main/LICENSE-DATA).

**How to cite.** Subash, Ambika. 2026. *Who owns India's plant varieties? PVP
Certification Explorer (India).* Centre for Economic Studies and Planning, Jawaharlal
Nehru University. <https://doi.org/10.5281/zenodo.21140353>

<details><summary style="cursor:pointer;display:inline-block;padding:2px 10px;margin-top:0.25rem;background:#eef2f7;border-radius:4px;font-weight:600;color:#374151;font-size:0.78rem">Show BibTeX</summary>

```
@misc{subash2026pvp,
  author      = {Subash, Ambika},
  title       = {Who owns {India}'s plant varieties? {PVP} Certification Explorer ({India})},
  year        = {2026},
  institution = {Centre for Economic Studies and Planning, Jawaharlal Nehru University},
  doi         = {10.5281/zenodo.21140353},
  url         = {https://doi.org/10.5281/zenodo.21140353},
  note        = {Data updated weekly. Accessed: [date]}
}
```

</details>

**About the author.** Built by Ambika Subash, Centre for Economic Studies and Planning,
Jawaharlal Nehru University. Her research examines the political economy of agricultural
biotechnology regulation, seed systems, and regulatory governance in India.
[Substack](https://theloneresearcher.substack.com/) |
[ORCID](https://orcid.org/0000-0002-8636-7123) |
[Google Scholar](https://scholar.google.com/citations?user=QbSSDTAAAAAJ&hl=en)
:::
