---
title: "Full analysis"
subtitle: "Certificates at a glance, the headline findings, and the cotton case study"
---

```{r setup}
source("R/prep_dashboard.R")
```

::: {.callout-note appearance="minimal"}
**`r format(n_total, big.mark=",")` certificates** analysed, from the
[PPV&FR Authority register](https://plantauthority.gov.in/list-certificates).
Last data refresh: **`r last_updated`**. This page rebuilds every week.
:::

The Protection of Plant Varieties and Farmers' Rights Act (PPVFRA), 2001 was designed as a
sui generis intellectual property regime recognising three classes of right-holders:
breeders, researchers, and farmers. The certification record covers every plant variety
protection certificate issued since 2007. This page asks who has actually benefited from
the Act's operation. Cotton receives separate treatment throughout: it is India's only
commercially cultivated genetically modified crop, and the site where private
concentration is most complete.

To ask these questions of a specific crop yourself, use the
[interactive explorer](index.qmd).

# Certificates at a glance

## Certificates issued over time

```{r}
f1 <- dd %>% count(issue_year)
year_2013_n <- first(f1$n[f1$issue_year == 2013], default = NA)
year_2022_n <- first(f1$n[f1$issue_year == 2022], default = NA)

plot_ly(f1, x = ~issue_year, y = ~n, type = "scatter", mode = "lines+markers",
        line = list(color = accent), marker = list(color = accent),
        hovertemplate = "%{x}: %{y} certificates<extra></extra>") %>%
  layout(title = "Total PPVFRA certificates issued, all crops",
         xaxis = list(title = "Year of certificate issue"),
         yaxis = list(title = "No. of certificates", tickformat = ",d"),
         annotations = list(
           list(x = 2013, y = year_2013_n, text = "2013", showarrow = TRUE, arrowhead = 2, ay = -35),
           list(x = 2022, y = year_2022_n, text = "2022", showarrow = TRUE, arrowhead = 2, ay = -35)))
```

Certification grew slowly in the Act's early years, peaked in 2013-14 due to a surge in
NGO-facilitated farmer variety registrations, then fell during the Covid years before
recovering to its highest-ever levels in 2023-24. The post-2022 recovery is dominated by
private applicants, particularly in cotton.

## Cotton, on its own

```{r}
f2 <- cot %>% count(issue_year)
plot_ly(f2, x = ~issue_year, y = ~n, type = "scatter", mode = "lines+markers",
        line = list(color = accent), marker = list(color = accent),
        hovertemplate = "%{x}: %{y} certificates<extra></extra>") %>%
  layout(title = "PPVFRA certificates issued, cotton",
         xaxis = list(title = "Year of certificate issue"),
         yaxis = list(title = "No. of certificates", tickformat = ",d"))
```

Cotton variety certification has accelerated sharply since 2022, reaching its highest-ever
annual total in 2024. Private firms account for virtually all of this growth. SAU
registrations, which were active until around 2014, have not recovered.

## What gets certified

```{r}
f3 <- d %>% count(crop_bucket) %>%
  mutate(crop_bucket = fct_relevel(fct_reorder(crop_bucket, n, .desc = TRUE), "Others", after = Inf))
ggplot(f3, aes(crop_bucket, n)) +
  geom_col(fill = accent) + geom_text(aes(label = comma(n)), vjust = -0.4, size = 3) +
  scale_y_continuous(labels = comma) +
  labs(title = "Certificates by crop group", x = NULL, y = "Certificates issued") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

Cereals dominate, driven by rice registrations. Vegetables are the second-largest group,
ahead of fibre crops (which are entirely cotton). Cotton's outsized influence on the pages
that follow reflects its commercial significance and the intensity of private breeding
activity concentrated in a single crop, not its raw certificate count.

## By variety category

```{r}
f4 <- d %>% filter(!is.na(variety_category)) %>% count(variety_category)
ggplot(f4, aes(variety_category, n, fill = variety_category)) +
  geom_col() + scale_fill_manual(values = cat_cols) + guides(fill = "none") +
  geom_text(aes(label = comma(n)), vjust = -0.4, size = 3) +
  scale_y_continuous(labels = comma) +
  labs(title = "Certificates by variety category", x = NULL, y = "Certificates issued")
```

Farmers' varieties are the largest single category in aggregate, but this number requires
qualification. Farmer registrations are concentrated in traditional food crops, confer no
right to sell branded seed, and generate no royalties. In commercially significant crops,
particularly cotton, the distribution inverts. Essentially Derived Varieties (EDVs), the
category designed to trigger benefit-sharing when registered varieties are used as the
basis for derivative breeding, account for just 14 certificates across the entire dataset.
All 14 are in cotton. All 14 are privately held.

## Farmers' varieties over time

```{r}
f5 <- dd %>% filter(variety_category == "Farmer") %>% count(issue_year)
ggplot(f5, aes(issue_year, n)) +
  geom_col(fill = cat_cols[["Farmer"]]) + geom_text(aes(label = n), vjust = -0.4, size = 3) +
  scale_x_continuous(breaks = seq(yr_min, max(f5$issue_year), 1)) +
  scale_y_continuous(labels = comma) +
  labs(title = "Certificates issued for farmers' varieties annually",
       x = "Year of certificate issue", y = "Farmers' certificates") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Farmer variety certification peaked between 2012 and 2015, driven by NGO-facilitated bulk
registrations of traditional landraces. Since 2016 it has declined sharply, with near-zero
issuances in recent years. No farmer has received a certificate for a new variety under
the Act.

## Public applicants, by category

```{r}
f6 <- dd %>% filter(applicant_category == "Public", !is.na(variety_category)) %>% count(issue_year, variety_category)
ggplot(f6, aes(issue_year, n, fill = variety_category)) + geom_col() +
  scale_fill_manual(values = cat_cols) +
  scale_x_continuous(breaks = seq(yr_min, max(f6$issue_year), 2)) +
  scale_y_continuous(labels = comma) +
  labs(title = "Certificates to public applicants, by category",
       x = "Year of certificate issue", y = "Certificates", fill = "Category")
```

Public institution registrations are concentrated in extant and notified variety
categories, reflecting existing varieties registered for protection rather than newly
developed commercial innovations. ICAR's large registration count reflects Green
Revolution-era breeding momentum rather than contemporary activity. New variety
registrations by public institutions have declined significantly since the mid-2010s.

## Private applicants, by category

```{r}
f7 <- dd %>% filter(applicant_category == "Private", !is.na(variety_category)) %>% count(issue_year, variety_category)
ggplot(f7, aes(issue_year, n, fill = variety_category)) + geom_col() +
  scale_fill_manual(values = cat_cols) +
  scale_x_continuous(breaks = seq(yr_min, max(f7$issue_year), 2)) +
  scale_y_continuous(labels = comma) +
  labs(title = "Certificates to private applicants, by category",
       x = "Year of certificate issue", y = "Certificates", fill = "Category")
```

Private applicants register predominantly new varieties and Varieties of Common Knowledge
(VCKs). The VCK category is significant: these are varieties already in the public domain
that are converted into private assets through registration. Of 293 cotton VCKs in the
dataset, 292 are privately held and 1 is held by a public institution.

## Ownership concentration beyond cotton

```{r}
top_firms_bar(d, title = "Major private applicants, all crops")
```

Market concentration is high across all crops, not just cotton. A small number of
companies dominate private variety certification in India. Cotton, covered next, is the
most extreme version of this pattern.

# The headline findings

::: {.callout-note appearance="simple"}
These charts carry the argument of this page. A reader who only looks at this section
should leave knowing the story.
:::

## Applicant category, cotton vs all crops

```{r}
f8 <- bind_rows(d %>% count(sector) %>% mutate(set = "All crops"),
                cot %>% count(sector) %>% mutate(set = "Cotton")) %>% filter(!is.na(sector))
ggplot(f8, aes(sector, n, fill = set)) +
  geom_col(position = "dodge") + scale_fill_manual(values = set_cols) +
  geom_text(aes(label = comma(n)), position = position_dodge(width = 0.9), vjust = -0.4, size = 3) +
  scale_y_continuous(labels = comma) +
  labs(title = "Applicant category: cotton vs all crops",
       x = "Applicant category", y = "Certificates", fill = NULL)
```

Across all crops, farmers hold the largest share of certificates, driven by
NGO-facilitated registrations of traditional varieties in food crops. In cotton the
distribution inverts. Private firms hold 585 of 696 cotton certificates. Farmers hold 1.
Transgenic hybrid production requires trait licensing agreements, backcross breeding
infrastructure, and biosafety containment facilities. These are not administrative
barriers reducible to policy adjustment; they are the technical conditions of production
in commercial cotton, and they are inaccessible to farmers as a class.

## Cumulative ownership over time, all crops

```{r}
f_cum <- dd %>% filter(!is.na(sector)) %>% count(issue_year, sector) %>%
  arrange(issue_year) %>% group_by(sector) %>% mutate(cum = cumsum(n)) %>% ungroup()
ggplot(f_cum, aes(issue_year, cum, color = sector)) +
  geom_line(linewidth = 1.1) + geom_point(size = 1.5) +
  scale_color_manual(values = sector_cols) +
  scale_x_continuous(breaks = seq(yr_min, max(f_cum$issue_year), 2)) +
  scale_y_continuous(labels = comma) +
  labs(title = "Cumulative certificates by applicant category, all crops",
       x = "Year of certificate issue", y = "Cumulative certificates", color = NULL)
```

This is the inversion in the previous chart, unfolding year by year rather than as a single
endpoint. Watch where the lines separate: that is the moment the structural divergence
between applicant types took hold.

## Cotton applicant composition over time

```{r}
f10 <- cot %>% filter(!is.na(sector)) %>% count(issue_year, sector)
plot_ly(f10, x = ~issue_year, y = ~n, color = ~sector, colors = sector_cols,
        type = "scatter", mode = "lines+markers") %>%
  layout(title = "Cotton applicant composition over time",
         xaxis = list(title = "Year of certificate issue"),
         yaxis = list(title = "No. of certificates", tickformat = ",d"))
```

Private applicants do not just dominate cotton certification, they have driven almost
every year of its growth. Public, SAU, and farmer lines stay close to flat by comparison.

## Who are the private cotton firms?

```{r}
top_firms_bar(cot, title = "Major private applicants, cotton")
```

In cotton, five firms account for more than 60 percent of all private certificates issued
between 2009 and 2024: Nuziveedu Seeds (30.1 percent), Prabhat Agri Biotech (11.8 percent),
Mahyco (7.1 percent), Kaveri Seed Co (6.6 percent), and Rasi Seeds (6.1 percent). These
figures cover direct variety registrations only. Trait-level control through
sublicensing, particularly Monsanto/Bayer's Bt and BGII events embedded across virtually
all tetraploid registrations, is not visible in registration data and is not captured
here.

## How concentrated is private cotton ownership?

```{r}
priv_cot_firms <- cot %>% filter(applicant_category == "Private") %>%
  mutate(key = alias_company(canon_company(applicant))) %>%
  count(key, name = "n") %>% arrange(n)
lorenz_df <- priv_cot_firms %>%
  mutate(cum_firms = row_number() / n(), cum_certs = cumsum(n) / sum(n)) %>%
  select(cum_firms, cum_certs) %>%
  bind_rows(tibble(cum_firms = 0, cum_certs = 0), .)
gini_cotton <- gini(priv_cot_firms$n)

ggplot(lorenz_df, aes(cum_firms, cum_certs)) +
  geom_area(fill = sector_cols[["Private"]], alpha = 0.15) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = muted) +
  geom_line(color = sector_cols[["Private"]], linewidth = 1.1) +
  scale_x_continuous(labels = percent, limits = c(0, 1)) +
  scale_y_continuous(labels = percent, limits = c(0, 1)) +
  labs(title = "Concentration of private cotton certificates among firms",
       x = "Share of firms, smallest to largest", y = "Cumulative share of certificates")
```

The dashed line is perfect equality, where every firm holds the same share. The further
the solid line bows away from it, the more concentrated ownership is. The Gini
coefficient here is **`r round(gini_cotton, 2)`**, on a scale from 0 (every firm holds an
equal share) to 1 (one firm holds everything).

## Public vs private innovation in cotton

```{r}
f_new_pubpriv <- cot %>% filter(variety_category == "New", sector %in% c("Public", "Private")) %>%
  count(issue_year, sector)
ggplot(f_new_pubpriv, aes(issue_year, n, color = sector)) +
  geom_line(linewidth = 1.1) + geom_point(size = 1.5) +
  scale_color_manual(values = sector_cols) +
  scale_x_continuous(breaks = seq(yr_min, max(f_new_pubpriv$issue_year, na.rm = TRUE), 2)) +
  scale_y_continuous(labels = comma) +
  labs(title = "New cotton variety registrations, public vs private",
       x = "Year of certificate issue", y = "New variety certificates", color = NULL)
```

This excludes VCKs and extant varieties, leaving only registrations of genuinely new
material. It isolates active breeding from the registration of already-existing
varieties, which is a separate process covered next.

# Cotton in detail

The headline findings established that cotton inverts the usual applicant pattern. This
section supports that finding with more detail on how.

## Ploidy by applicant

```{r}
f9 <- cot %>% filter(!is.na(ploidy), !is.na(sector)) %>% count(sector, ploidy)
ggplot(f9, aes(sector, n, fill = ploidy)) +
  geom_col(position = "dodge") + scale_fill_manual(values = ploidy_cols) +
  geom_text(aes(label = n), position = position_dodge(width = 0.9), vjust = -0.4, size = 3) +
  scale_y_continuous(labels = comma) +
  labs(title = "Cotton certificates by ploidy", x = "Applicant category",
       y = "No. of PVP certificates", fill = NULL)
```

Private cotton certification is almost entirely tetraploid, the hybrid and Bt lines
preferred by the textile industry for their longer staple length. Diploid or desi cotton,
historically adapted to low-input smallholder conditions, has been abandoned by both
private and public breeders. SAUs show a more balanced ploidy profile, reflecting a
residual public mandate to maintain crop diversity, but SAU cotton registrations have
effectively ceased since 2014.

```{r}
farmer_cot <- cot %>% filter(sector == "Farmer") %>%
  select(issue_year, variety_category, crop, applicant)
```

::: {.callout-note appearance="simple"}
**The farmer presence in cotton, in full.** `r nrow(farmer_cot)` certificate(s) out of
`r comma(nrow(cot))` cotton certificates were issued to a farmer applicant.
`r if (nrow(farmer_cot) >= 1) paste0("The first was issued in ", farmer_cot$issue_year[1], ", category ", as.character(farmer_cot$variety_category[1]), ", crop ", farmer_cot$crop[1], ".") else ""`
:::

## Variety category by applicant

```{r}
f11 <- cot %>% filter(!is.na(variety_category), !is.na(sector)) %>% count(variety_category, sector)
ggplot(f11, aes(variety_category, n, fill = sector)) + geom_col() +
  scale_fill_manual(values = sector_cols) +
  geom_text(aes(label = n), position = position_stack(vjust = 0.5), size = 3) +
  scale_y_continuous(labels = comma) +
  labs(title = "Cotton certificates by variety category and applicant",
       x = "Variety category", y = "Certificates", fill = NULL) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

Private firms register primarily new varieties and VCKs. Public institutions register
mostly extant and notified varieties. Farmers are confined to the farmers' variety
category. EDVs appear only in the private column. Only 14 EDV certificates have been
issued nationally across 17 years of the Act's operation, all in cotton. Whether firms
avoid the EDV category because of its benefit-sharing obligations, or register derivative
varieties as new varieties by meeting the DUS distinctness threshold through minor
morphological modification, the registration data cannot resolve.

## Every EDV certificate ever issued

```{r}
f_edv <- d %>% filter(variety_category == "EDV") %>% count(issue_year)
edv_total  <- sum(d$variety_category == "EDV", na.rm = TRUE)
edv_crops  <- paste(sort(unique(as.character(d$crop_bucket[d$variety_category == "EDV"]))), collapse = ", ")
edv_years  <- sort(unique(f_edv$issue_year))
edv_yr_txt <- if (length(edv_years) == 1) as.character(edv_years) else paste(range(edv_years), collapse = "-")

edv_axis_max <- max(d$issue_year, na.rm = TRUE)
ggplot(f_edv, aes(issue_year, n)) +
  geom_segment(aes(xend = issue_year, y = 0, yend = n), color = cat_cols[["EDV"]], linewidth = 1) +
  geom_point(size = 4, color = cat_cols[["EDV"]]) +
  geom_text(aes(label = n), vjust = -1.2, size = 3.5) +
  scale_x_continuous(breaks = seq(yr_min, edv_axis_max, 1),
                      limits = c(yr_min - 1, edv_axis_max + 1)) +
  scale_y_continuous(limits = c(0, max(f_edv$n) * 1.3), breaks = pretty_breaks()) +
  labs(title = "Essentially Derived Variety (EDV) certificates, all years, all crops",
       x = "Year of certificate issue", y = "EDV certificates issued") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

`r edv_total` EDV certificates have been issued nationally, all in `r edv_crops`,
concentrated in `r edv_yr_txt`. No EDV has been registered before or after that window, in
any crop.

## Testing the VCK enclosure sequence

```{r}
f_vck_new <- cot %>% filter(applicant_category == "Private", variety_category %in% c("New", "Extant (VCK)")) %>%
  count(issue_year, variety_category)
ggplot(f_vck_new, aes(issue_year, n, fill = variety_category)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = cat_cols) +
  scale_x_continuous(breaks = seq(yr_min, max(f_vck_new$issue_year, na.rm = TRUE), 2)) +
  scale_y_continuous(labels = comma) +
  labs(title = "Private cotton: VCK vs New variety registrations over time",
       x = "Year of certificate issue", y = "Certificates", fill = "Category")
```

One reading of the VCK category is a sequential strategy: register existing
common-knowledge varieties first, then shift to new variety development once that
low-cost registration opportunity is exhausted.

## Private cotton, by category over time

```{r}
f12 <- cot %>% filter(applicant_category == "Private", !is.na(variety_category)) %>% count(issue_year, variety_category)
ggplot(f12, aes(issue_year, n, fill = variety_category)) + geom_col() +
  scale_fill_manual(values = cat_cols) +
  geom_text(aes(label = n), position = position_stack(vjust = 0.5), size = 3) +
  scale_x_continuous(breaks = seq(yr_min, max(f12$issue_year), 2)) +
  scale_y_continuous(labels = comma) +
  labs(title = "Private cotton certificates by category",
       x = "Year of certificate issue", y = "Certificates", fill = "Category")
```

Private cotton certification shifted over time from VCK registrations toward new variety
registrations, particularly after 2016. The 2024 surge is almost entirely new varieties.
EDV registrations appeared briefly in 2019 and have not recurred.

---

::: {.callout-tip appearance="minimal"}
Data source: Protection of Plant Varieties and Farmers' Rights Authority public
certificate register, scraped weekly. Applicant categories are normalised from raw
register entries. All figures use year of certificate issue. Full pipeline is open source
and available in the
[project repository](https://github.com/ambika-subash/pvp-certification-explorer).
Built with R and Quarto.
:::
