---
title: "Who owns India's plant varieties?"
subtitle: "Evidence from PPV&FRA certification data"
---

```{r setup}
library(dplyr); library(readr); library(tidyr); library(stringr)
library(ggplot2); library(forcats); library(scales); library(plotly)

d <- read_csv("data/pvp_certificates_clean.csv", show_col_types = FALSE)

yr_min <- 2009L
theme_set(theme_minimal(base_size = 13))
blue <- "#3b6fb6"
cat_cols <- c("Farmer"="#4daf4a","New"="#377eb8","Extant"="#80b1d3",
              "Extant (Notified)"="#ff7f00","Extant (VCK)"="#e41a1c","EDV"="#984ea3")
sector_cols <- c("Farmer"="#4daf4a","Private"="#377eb8","Public"="#e41a1c",
                 "SAU"="#ff7f00","Individual Breeder"="#984ea3")
ploidy_cols <- c("Diploid"="#377eb8","Tetraploid"="#e41a1c")
set_cols    <- c("All crops"="#377eb8","Cotton"="#e41a1c")
sector_levels <- c("Farmer","Private","Public","SAU","Individual Breeder")
cat_levels    <- c("Farmer","New","Extant","Extant (Notified)","Extant (VCK)","EDV")
paper_groups  <- c("Cereals","Fibre Crops","Vegetables","Legumes","Oilseeds","Fruits","Sugar Crops")

d <- d %>% mutate(
  crop_bucket      = ifelse(crop_group %in% paper_groups, crop_group, "Others"),
  sector           = factor(sector, levels = sector_levels),
  variety_category = factor(variety_category, levels = cat_levels))
dd  <- d %>% filter(!is.na(issue_year), issue_year >= yr_min)
cot <- dd %>% filter(is_cotton) %>%
  mutate(ploidy = case_when(
    str_detect(crop, regex("tetraploid", ignore_case = TRUE)) ~ "Tetraploid",
    str_detect(crop, regex("diploid",    ignore_case = TRUE)) ~ "Diploid",
    TRUE ~ NA_character_))

canon_company <- function(x){
  y <- str_to_lower(x); y <- str_replace_all(y, "[.,()]", " ")
  y <- str_replace_all(y, "&", " and "); y <- str_replace_all(y, "\\bseeds\\b", "seed")
  y <- str_replace_all(y, "\\b(private|pvt|limited|ltd|llp|company|co|corporation|corp|incorporated|inc)\\b", " ")
  str_squish(y)
}
alias_company <- function(key) case_when(
  str_detect(key, "maharashtra hybrid seed") | key == "mahyco" ~ "mahyco",
  str_detect(key, "^monsanto") ~ "monsanto", TRUE ~ key)
pretty_names <- c("mahyco" = "Mahyco (Maharashtra Hybrid Seeds)", "monsanto" = "Monsanto")

n_total <- nrow(d)
last_updated <- tryCatch(sub("^Last updated: ", "", readLines("data/last_updated.txt", warn = FALSE)[1]),
                         error = function(e) format(Sys.Date()))
```

::: {.callout-note appearance="minimal"}
**`r format(n_total, big.mark=",")` certificates** analysed, from the
[PPV&FR Authority register](https://plantauthority.gov.in/list-certificates).
Last data refresh: **`r last_updated`**. This page rebuilds every week.
:::

The Protection of Plant Varieties and Farmers' Rights Act (PPVFRA), 2001 set out to
protect breeders' rights while also recognising farmers as innovators. This page reports
what the certification record shows: who secures plant variety protection, for which
crops and categories, and how that has changed over time. Cotton gets separate attention
throughout, since it is India's only widely genetically modified crop.

# Certificates over time

## All crops

```{r}
f1 <- dd %>% count(issue_year)
plot_ly(f1, x = ~issue_year, y = ~n, type = "scatter", mode = "lines+markers",
        line = list(color = blue), marker = list(color = blue),
        hovertemplate = "%{x}: %{y} certificates<extra></extra>") %>%
  layout(title = "Total PPVFRA certificates issued, all crops",
         xaxis = list(title = "Year of certificate issue"),
         yaxis = list(title = "No. of certificates"))
```

Certification started slowly, fell during the Covid years, and has since climbed to its
highest levels on record.

## Cotton

```{r}
f2 <- cot %>% count(issue_year)
plot_ly(f2, x = ~issue_year, y = ~n, type = "scatter", mode = "lines+markers",
        line = list(color = blue), marker = list(color = blue),
        hovertemplate = "%{x}: %{y} certificates<extra></extra>") %>%
  layout(title = "PPVFRA certificates issued, cotton",
         xaxis = list(title = "Year of certificate issue"),
         yaxis = list(title = "No. of certificates"))
```

# What gets certified

## By crop group

```{r}
f3 <- d %>% count(crop_bucket) %>%
  mutate(crop_bucket = fct_relevel(fct_reorder(crop_bucket, n, .desc = TRUE), "Others", after = Inf))
ggplot(f3, aes(crop_bucket, n)) +
  geom_col(fill = blue) + geom_text(aes(label = n), vjust = -0.4, size = 3) +
  labs(title = "Certificates by crop group", x = NULL, y = "Certificates issued") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

Cereals lead by a wide margin. Vegetables are the second-largest group, followed by
fibre crops (which are almost all cotton).

## By variety category

```{r}
f4 <- d %>% filter(!is.na(variety_category)) %>% count(variety_category)
ggplot(f4, aes(variety_category, n, fill = variety_category)) +
  geom_col() + scale_fill_manual(values = cat_cols) + guides(fill = "none") +
  geom_text(aes(label = n), vjust = -0.4, size = 3) +
  labs(title = "Certificates by variety category", x = NULL, y = "Certificates issued")
```

Farmers' varieties are the largest single category overall. The crop by crop breakdown
below shows that this total sits very differently across crops.

# Public, private, and farmers

Certificates are filed by five types of applicants: Farmer (an individual farmer or farmers' group), Private (for-profit firms), Public (government research bodies), SAU (State Agricultural Universities), and Individual Breeder (independent plant breeders not affiliated with any organization).

## Applicant category, cotton vs all crops

```{r}
f8 <- bind_rows(d %>% count(sector) %>% mutate(set = "All crops"),
                cot %>% count(sector) %>% mutate(set = "Cotton")) %>% filter(!is.na(sector))
ggplot(f8, aes(sector, n, fill = set)) +
  geom_col(position = "dodge") + scale_fill_manual(values = set_cols) +
  geom_text(aes(label = n), position = position_dodge(width = 0.9), vjust = -0.4, size = 3) +
  labs(title = "Applicant category: cotton vs all crops",
       x = "Applicant category", y = "Certificates", fill = NULL)
```

Farmers hold the most certificates across all crops. In cotton the order reverses:
private firms hold almost all of them, and farmers hold close to none.

## Farmers' varieties over time

```{r}
f5 <- dd %>% filter(variety_category == "Farmer") %>% count(issue_year)
ggplot(f5, aes(issue_year, n)) +
  geom_col(fill = cat_cols[["Farmer"]]) + geom_text(aes(label = n), vjust = -0.4, size = 3) +
  scale_x_continuous(breaks = seq(yr_min, max(f5$issue_year), 1)) +
  labs(title = "Certificates issued for farmers' varieties annually",
       x = "Year of certificate issue", y = "Farmers' certificates") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Public applicants, by category

```{r}
f6 <- dd %>% filter(applicant_category == "Public", !is.na(variety_category)) %>% count(issue_year, variety_category)
ggplot(f6, aes(issue_year, n, fill = variety_category)) + geom_col() +
  scale_fill_manual(values = cat_cols) +
  scale_x_continuous(breaks = seq(yr_min, max(f6$issue_year), 2)) +
  labs(title = "Certificates to public applicants, by category",
       x = "Year of certificate issue", y = "Certificates", fill = "Category")
```

## Private applicants, by category

```{r}
f7 <- dd %>% filter(applicant_category == "Private", !is.na(variety_category)) %>% count(issue_year, variety_category)
ggplot(f7, aes(issue_year, n, fill = variety_category)) + geom_col() +
  scale_fill_manual(values = cat_cols) +
  scale_x_continuous(breaks = seq(yr_min, max(f7$issue_year), 2)) +
  labs(title = "Certificates to private applicants, by category",
       x = "Year of certificate issue", y = "Certificates", fill = "Category")
```
## Company concentration across all crops

```{r}
priv_all <- d %>% filter(applicant_category == "Private") %>%
  mutate(key = alias_company(canon_company(applicant)))
labels_all <- priv_all %>% count(key, applicant) %>% group_by(key) %>%
  slice_max(n, n = 1, with_ties = FALSE) %>% ungroup() %>%
  transmute(key, label = coalesce(unname(pretty_names[key]), applicant))
f14 <- priv_all %>% count(key, name = "n") %>% left_join(labels_all, by = "key") %>%
  arrange(desc(n)) %>%
  mutate(label = ifelse(row_number() <= 10, label, "Others")) %>%
  group_by(label) %>% summarise(n = sum(n), .groups = "drop") %>%
  mutate(pct = n / sum(n)) %>% arrange(desc(n))
ggplot(f14, aes("", n, fill = fct_reorder(label, n, .desc = TRUE))) +
  geom_col(width = 1, color = "white") + coord_polar(theta = "y") +
  geom_text(aes(label = ifelse(pct >= 0.03, percent(pct, accuracy = 0.1), "")),
            position = position_stack(vjust = 0.5), size = 3) +
  labs(title = "Major private applicants, all crops", fill = NULL) +
  theme_void(base_size = 12)

# The cotton case

## Who are the private cotton firms?

```{r}
priv <- cot %>% filter(applicant_category == "Private") %>%
  mutate(key = alias_company(canon_company(applicant)))
labels <- priv %>% count(key, applicant) %>% group_by(key) %>%
  slice_max(n, n = 1, with_ties = FALSE) %>% ungroup() %>%
  transmute(key, label = coalesce(unname(pretty_names[key]), applicant))
f13 <- priv %>% count(key, name = "n") %>% left_join(labels, by = "key") %>%
  arrange(desc(n)) %>%
  mutate(label = ifelse(row_number() <= 10, label, "Others")) %>%
  group_by(label) %>% summarise(n = sum(n), .groups = "drop") %>%
  mutate(pct = n / sum(n)) %>% arrange(desc(n))
ggplot(f13, aes("", n, fill = fct_reorder(label, n, .desc = TRUE))) +
  geom_col(width = 1, color = "white") + coord_polar(theta = "y") +
  geom_text(aes(label = ifelse(pct >= 0.03, percent(pct, accuracy = 0.1), "")),
            position = position_stack(vjust = 0.5), size = 3) +
  labs(title = "Major private applicants, cotton", fill = NULL) +
  theme_void(base_size = 12)
```

A small number of firms, led by Nuziveedu, hold most of the private cotton certificates.
Varietal control in India's most commercial crop rests with a few companies.

---

## Ploidy by applicant

```{r}
f9 <- cot %>% filter(!is.na(ploidy), !is.na(sector)) %>% count(sector, ploidy)
ggplot(f9, aes(sector, n, fill = ploidy)) +
  geom_col(position = "dodge") + scale_fill_manual(values = ploidy_cols) +
  geom_text(aes(label = n), position = position_dodge(width = 0.9), vjust = -0.4, size = 3) +
  labs(title = "Cotton certificates by ploidy", x = "Applicant category",
       y = "No. of PVP certificates", fill = NULL)
```

Private cotton certification is mostly tetraploid, the hybrid and Bt lines. Diploid desi
cotton makes up a small share for every type of applicant.

## Applicant composition over time

```{r}
f10 <- cot %>% filter(!is.na(sector)) %>% count(issue_year, sector)
plot_ly(f10, x = ~issue_year, y = ~n, color = ~sector, colors = sector_cols,
        type = "scatter", mode = "lines+markers") %>%
  layout(title = "Cotton applicant composition over time",
         xaxis = list(title = "Year of certificate issue"),
         yaxis = list(title = "No. of certificates"))
```

## Variety category by applicant

```{r}
f11 <- cot %>% filter(!is.na(variety_category), !is.na(sector)) %>% count(variety_category, sector)
ggplot(f11, aes(variety_category, n, fill = sector)) + geom_col() +
  scale_fill_manual(values = sector_cols) +
  labs(title = "Cotton certificates by variety category and applicant",
       x = "Variety category", y = "Certificates", fill = NULL) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

## Private cotton, by category over time

```{r}
f12 <- cot %>% filter(applicant_category == "Private", !is.na(variety_category)) %>% count(issue_year, variety_category)
ggplot(f12, aes(issue_year, n, fill = variety_category)) + geom_col() +
  scale_fill_manual(values = cat_cols) +
  scale_x_continuous(breaks = seq(yr_min, max(f12$issue_year), 2)) +
  labs(title = "Private cotton certificates by category",
       x = "Year of certificate issue", y = "Certificates", fill = "Category")
```


::: {.callout-tip appearance="minimal"}
Full pipeline (scrape, clean, figures, and this page) is in the
[project repository](https://github.com/ambika-subash/pvp-certification-explorer).
:::
