---
title: "Who owns India's plant varieties?"
subtitle: "Evidence from PPV&FRA certification data"
---

```{r setup}
library(dplyr); library(readr); library(tidyr); library(stringr)
library(ggplot2); library(forcats); library(scales); library(plotly)

d <- read_csv("data/pvp_certificates_clean.csv", show_col_types = FALSE)

yr_min <- 2009L
theme_set(theme_minimal(base_size = 13))
neutral <- "#6b7280"
cat_cols <- c("Farmer"="#4daf4a","New"="#377eb8","Extant"="#80b1d3",
              "Extant (Notified)"="#ff7f00","Extant (VCK)"="#e41a1c","EDV"="#984ea3")
sector_cols <- c("Farmer"="#4daf4a","Private"="#377eb8","Public"="#e41a1c",
                 "SAU"="#ff7f00","Individual Breeder"="#984ea3")
ploidy_cols <- c("Diploid"="#377eb8","Tetraploid"="#e41a1c")
set_cols    <- c("All crops"=neutral,"Cotton"="#e41a1c")
sector_levels <- c("Farmer","Private","Public","SAU","Individual Breeder")
cat_levels    <- c("Farmer","New","Extant","Extant (Notified)","Extant (VCK)","EDV")
paper_groups  <- c("Cereals","Fibre Crops","Vegetables","Legumes","Oilseeds","Fruits","Sugar Crops")

d <- d %>% mutate(
  crop_bucket      = ifelse(crop_group %in% paper_groups, crop_group, "Others"),
  sector           = factor(sector, levels = sector_levels),
  variety_category = factor(variety_category, levels = cat_levels))
dd  <- d %>% filter(!is.na(issue_year), issue_year >= yr_min)
cot <- dd %>% filter(is_cotton) %>%
  mutate(ploidy = case_when(
    str_detect(crop, regex("tetraploid", ignore_case = TRUE)) ~ "Tetraploid",
    str_detect(crop, regex("diploid",    ignore_case = TRUE)) ~ "Diploid",
    TRUE ~ NA_character_))

canon_company <- function(x){
  y <- str_to_lower(x); y <- str_replace_all(y, "[.,()]", " ")
  y <- str_replace_all(y, "&", " and "); y <- str_replace_all(y, "\\bseeds\\b", "seed")
  y <- str_replace_all(y, "\\b(private|pvt|limited|ltd|llp|company|co|corporation|corp|incorporated|inc)\\b", " ")
  str_squish(y)
}
alias_company <- function(key) case_when(
  str_detect(key, "maharashtra hybrid seed") | key == "mahyco" ~ "mahyco",
  str_detect(key, "^monsanto") ~ "monsanto", TRUE ~ key)
pretty_names <- c("mahyco" = "Mahyco (Maharashtra Hybrid Seeds)", "monsanto" = "Monsanto")

top_firms_bar <- function(df, n_top = 8, title){
  priv <- df %>% filter(applicant_category == "Private") %>%
    mutate(key = alias_company(canon_company(applicant)))
  labels <- priv %>% count(key, applicant) %>% group_by(key) %>%
    slice_max(n, n = 1, with_ties = FALSE) %>% ungroup() %>%
    transmute(key, label = coalesce(unname(pretty_names[key]), applicant))
  f <- priv %>% count(key, name = "n") %>% left_join(labels, by = "key") %>%
    arrange(desc(n)) %>%
    mutate(label = ifelse(row_number() <= n_top, label, "Others")) %>%
    group_by(label) %>% summarise(n = sum(n), .groups = "drop") %>% arrange(n)
  ggplot(f, aes(x = n, y = fct_reorder(label, n))) +
    geom_col(fill = sector_cols[["Private"]]) +
    geom_text(aes(label = comma(n)), hjust = -0.2, size = 3) +
    scale_x_continuous(labels = comma, expand = expansion(mult = c(0, .18))) +
    labs(title = title, x = "Certificates", y = NULL)
}

n_total <- nrow(d)
last_updated <- tryCatch(sub("^Last updated: ", "", readLines("data/last_updated.txt", warn = FALSE)[1]),
                         error = function(e) format(Sys.Date()))
```

::: {.callout-note appearance="minimal"}
**`r format(n_total, big.mark=",")` certificates** analysed, from the
[PPV&FR Authority register](https://plantauthority.gov.in/list-certificates).
Last data refresh: **`r last_updated`**. This page rebuilds every week.
:::

The Protection of Plant Varieties and Farmers' Rights Act (PPVFRA), 2001 set out to
protect breeders' rights while also recognising farmers as innovators. This page uses the
full certification register to show who actually holds plant variety protection, for which
crops, and how that has shifted over time.
[Read the full methodology](https://github.com/ambika-subash/pvp-certification-explorer/blob/main/README.md).

# Certificates at a glance

## Certificates issued over time

```{r}
f1 <- dd %>% count(issue_year)
year_2013_n  <- first(f1$n[f1$issue_year == 2013], default = NA)
latest_year  <- max(f1$issue_year)
latest_year_n <- first(f1$n[f1$issue_year == latest_year], default = NA)

plot_ly(f1, x = ~issue_year, y = ~n, type = "scatter", mode = "lines+markers",
        line = list(color = neutral), marker = list(color = neutral),
        hovertemplate = "%{x}: %{y} certificates<extra></extra>") %>%
  layout(title = "Total PPVFRA certificates issued, all crops",
         xaxis = list(title = "Year of certificate issue"),
         yaxis = list(title = "No. of certificates", tickformat = ",d"),
         annotations = list(
           list(x = 2013, y = year_2013_n, text = "2013", showarrow = TRUE, arrowhead = 2, ay = -35),
           list(x = latest_year, y = latest_year_n, text = as.character(latest_year),
                showarrow = TRUE, arrowhead = 2, ay = -35)))
```

Certificates jump sharply in 2013 (`r comma(year_2013_n)` issued) and again in
`r latest_year` (`r comma(latest_year_n)`). The register does not record why. These look
more like batches of applications cleared together than steady organic growth, but that is
an inference, not a confirmed cause, so treat it as an open question.

## What gets certified

```{r}
f3 <- d %>% count(crop_bucket) %>%
  mutate(crop_bucket = fct_relevel(fct_reorder(crop_bucket, n, .desc = TRUE), "Others", after = Inf))
ggplot(f3, aes(crop_bucket, n)) +
  geom_col(fill = neutral) + geom_text(aes(label = comma(n)), vjust = -0.4, size = 3) +
  scale_y_continuous(labels = comma) +
  labs(title = "Certificates by crop group", x = NULL, y = "Certificates issued") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

Cereals lead by a wide margin. Vegetables are the second-largest group, followed by
fibre crops (which are almost all cotton).

## By variety category

```{r}
f4 <- d %>% filter(!is.na(variety_category)) %>% count(variety_category)
ggplot(f4, aes(variety_category, n, fill = variety_category)) +
  geom_col() + scale_fill_manual(values = cat_cols) + guides(fill = "none") +
  geom_text(aes(label = comma(n)), vjust = -0.4, size = 3) +
  scale_y_continuous(labels = comma) +
  labs(title = "Certificates by variety category", x = NULL, y = "Certificates issued")
```

Farmers' varieties are the largest single category overall. The sections below show that
this total sits very differently across applicant types and crops.

## Farmers' varieties over time

```{r}
f5 <- dd %>% filter(variety_category == "Farmer") %>% count(issue_year)
ggplot(f5, aes(issue_year, n)) +
  geom_col(fill = cat_cols[["Farmer"]]) + geom_text(aes(label = n), vjust = -0.4, size = 3) +
  scale_x_continuous(breaks = seq(yr_min, max(f5$issue_year), 1)) +
  scale_y_continuous(labels = comma) +
  labs(title = "Certificates issued for farmers' varieties annually",
       x = "Year of certificate issue", y = "Farmers' certificates") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Farmer variety certification has effectively ceased in recent years. The bars that once
built the largest single category on this page have trailed off to almost nothing.

## Public applicants, by category

```{r}
f6 <- dd %>% filter(applicant_category == "Public", !is.na(variety_category)) %>% count(issue_year, variety_category)
ggplot(f6, aes(issue_year, n, fill = variety_category)) + geom_col() +
  scale_fill_manual(values = cat_cols) +
  scale_x_continuous(breaks = seq(yr_min, max(f6$issue_year), 2)) +
  scale_y_continuous(labels = comma) +
  labs(title = "Certificates to public applicants, by category",
       x = "Year of certificate issue", y = "Certificates", fill = "Category")
```

## Private applicants, by category

```{r}
f7 <- dd %>% filter(applicant_category == "Private", !is.na(variety_category)) %>% count(issue_year, variety_category)
ggplot(f7, aes(issue_year, n, fill = variety_category)) + geom_col() +
  scale_fill_manual(values = cat_cols) +
  scale_x_continuous(breaks = seq(yr_min, max(f7$issue_year), 2)) +
  scale_y_continuous(labels = comma) +
  labs(title = "Certificates to private applicants, by category",
       x = "Year of certificate issue", y = "Certificates", fill = "Category")
```

## Ownership concentration beyond cotton

```{r}
top_firms_bar(d, title = "Major private applicants, all crops")
```

Market concentration is high across all crops, not just cotton. A small number of
companies dominate private variety certification in India. Cotton, covered next, is the
most extreme version of this pattern.

# The headline findings

::: {.callout-note appearance="simple"}
These four charts carry the argument of this page. A reader who only looks at this section
should leave knowing the story.
:::

Certificates are filed by five types of applicants: Farmer (an individual farmer or
farmers' group), Private (for-profit firms), Public (government research bodies), SAU
(State Agricultural Universities), and Individual Breeder (independent plant breeders not
affiliated with any organisation).

## Applicant category, cotton vs all crops

```{r}
f8 <- bind_rows(d %>% count(sector) %>% mutate(set = "All crops"),
                cot %>% count(sector) %>% mutate(set = "Cotton")) %>% filter(!is.na(sector))
ggplot(f8, aes(sector, n, fill = set)) +
  geom_col(position = "dodge") + scale_fill_manual(values = set_cols) +
  geom_text(aes(label = comma(n)), position = position_dodge(width = 0.9), vjust = -0.4, size = 3) +
  scale_y_continuous(labels = comma) +
  labs(title = "Applicant category: cotton vs all crops",
       x = "Applicant category", y = "Certificates", fill = NULL)
```

This is the central finding. Farmers hold the most certificates of any applicant type
across all crops combined. In cotton, that order inverts almost completely: private firms
hold nearly every certificate, and farmers are left with a handful. The applicant category
that dominates the register overall is nearly absent in India's most commercial crop.

## Cumulative ownership over time, all crops

```{r}
f_cum <- dd %>% filter(!is.na(sector)) %>% count(issue_year, sector) %>%
  arrange(issue_year) %>% group_by(sector) %>% mutate(cum = cumsum(n)) %>% ungroup()
ggplot(f_cum, aes(issue_year, cum, color = sector)) +
  geom_line(linewidth = 1.1) + geom_point(size = 1.5) +
  scale_color_manual(values = sector_cols) +
  scale_x_continuous(breaks = seq(yr_min, max(f_cum$issue_year), 2)) +
  scale_y_continuous(labels = comma) +
  labs(title = "Cumulative certificates by applicant category, all crops",
       x = "Year of certificate issue", y = "Cumulative certificates", color = NULL)
```

This is the inversion in the previous chart, unfolding year by year rather than as a single
endpoint. Watch where the lines separate: that is the moment the structural divergence
between applicant types took hold.

## Cotton applicant composition over time

```{r}
f10 <- cot %>% filter(!is.na(sector)) %>% count(issue_year, sector)
plot_ly(f10, x = ~issue_year, y = ~n, color = ~sector, colors = sector_cols,
        type = "scatter", mode = "lines+markers") %>%
  layout(title = "Cotton applicant composition over time",
         xaxis = list(title = "Year of certificate issue"),
         yaxis = list(title = "No. of certificates", tickformat = ",d"))
```

Private applicants do not just dominate cotton certification, they have driven almost
every year of its growth. Public, SAU, and farmer lines stay close to flat by comparison.

## Who are the private cotton firms?

```{r}
top_firms_bar(cot, title = "Major private applicants, cotton")
```

A small number of firms, led by Nuziveedu, hold most of the private cotton certificates.
Varietal control in India's most commercial crop rests with a few companies.

# Explore the data yourself

::: {.callout-important appearance="simple"}
**These charts are interactive.** Each one has two dropdowns side by side: pick a crop
category (like Cereals) or a specific crop (like Brinjal), and the chart redraws.
Whichever dropdown you touch last is the one that applies.
:::

The findings above are about cotton specifically. The tools below let you ask the same
questions of any crop in the register.

```{r}
#| label: prep-ojs
#| include: false

# Individual crops offered in the dropdown: the 20 most-certified crops by name.
top_crops <- d %>% filter(!is.na(crop)) %>% count(crop, sort = TRUE) %>%
  slice_head(n = 20) %>% pull(crop)
crop_categories <- sort(unique(na.omit(d$crop_bucket)))

# Each row is tagged under both its crop group and its individual crop name,
# so either dropdown can filter the same underlying dataset.
sel_for <- function(df) bind_rows(
  df %>% filter(!is.na(crop_bucket)) %>% mutate(sel = as.character(crop_bucket)),
  df %>% filter(crop %in% top_crops)  %>% mutate(sel = crop))

chr <- function(df) df %>% mutate(across(where(is.factor), as.character))

trends_df <- chr(sel_for(dd) %>% filter(!is.na(sector)) %>%
  count(sel, issue_year, sector, name = "n"))

comp_all <- d %>% filter(!is.na(sector)) %>% count(sector, name = "n") %>% mutate(sel = "All crops")
comp_df  <- chr(bind_rows(comp_all, sel_for(d) %>% filter(!is.na(sector)) %>% count(sel, sector, name = "n")))

firm_key    <- sel_for(d) %>% filter(applicant_category == "Private") %>%
  mutate(firm = alias_company(canon_company(applicant)))
firm_labels <- firm_key %>% count(firm, applicant) %>% group_by(firm) %>%
  slice_max(n, n = 1, with_ties = FALSE) %>% ungroup() %>%
  transmute(firm, label = coalesce(unname(pretty_names[firm]), applicant))
firm_df <- chr(firm_key %>% count(sel, firm, name = "n") %>%
  group_by(sel) %>% slice_max(n, n = 10, with_ties = FALSE) %>% ungroup() %>%
  left_join(firm_labels, by = "firm"))

var_df <- chr(sel_for(d) %>% filter(!is.na(variety_category), !is.na(sector)) %>%
  count(sel, sector, variety_category, name = "n"))

cum_sel <- sel_for(dd) %>% filter(!is.na(sector)) %>% count(sel, issue_year, sector, name = "n")
cum_all <- dd %>% filter(!is.na(sector)) %>% count(issue_year, sector, name = "n") %>%
  mutate(sel = "All crops")
cum_df  <- chr(bind_rows(cum_all, cum_sel) %>% arrange(sel, sector, issue_year) %>%
  group_by(sel, sector) %>% mutate(cum = cumsum(n)) %>% ungroup())

ojs_define(categories_ojs = crop_categories)
ojs_define(crops_ojs      = top_crops)
ojs_define(trends_ojs = trends_df)
ojs_define(comp_ojs   = comp_df)
ojs_define(firm_ojs   = firm_df)
ojs_define(var_ojs    = var_df)
ojs_define(cum_ojs    = cum_df)
```

```{ojs}
//| echo: false
categories = categories_ojs
crops = crops_ojs
sectorDomain = ["Farmer", "Private", "Public", "SAU", "Individual Breeder"]
sectorRange  = ["#4daf4a", "#377eb8", "#e41a1c", "#ff7f00", "#984ea3"]
catDomain = ["Farmer", "New", "Extant", "Extant (Notified)", "Extant (VCK)", "EDV"]
catRange  = ["#4daf4a", "#377eb8", "#80b1d3", "#ff7f00", "#e41a1c", "#984ea3"]

// Custom paired viewof: a category dropdown and a crop dropdown side by side.
// Whichever one fires last wins; its value becomes this input's value.
function pairedCropInput({catOptions = categories, cropOptions = crops,
                           catValue, cropValue, catLabel, cropLabel,
                           defaultMode = "crop"} = {}) {
  const catSel  = Inputs.select(catOptions,  {label: catLabel  ?? "Crop category", value: catValue  ?? catOptions[0]});
  const cropSel = Inputs.select(cropOptions, {label: cropLabel ?? "Specific crop", value: cropValue ?? cropOptions[0]});
  const box = html`<div style="display:flex;gap:2rem;flex-wrap:wrap;align-items:flex-end;margin-bottom:0.5rem">
    <div>${catSel}</div><div>${cropSel}</div></div>`;
  box.value = defaultMode === "crop" ? cropSel.value : catSel.value;
  function fire(mode) {
    box.value = mode === "category" ? catSel.value : cropSel.value;
    box.dispatchEvent(new CustomEvent("input", {bubbles: true}));
  }
  catSel.addEventListener("input",  (e) => { e.stopPropagation(); fire("category"); });
  cropSel.addEventListener("input", (e) => { e.stopPropagation(); fire("crop"); });
  return box;
}
```

## Certificate trends by crop

```{ojs}
//| echo: false
viewof sel1 = pairedCropInput({catValue: "Cereals", cropValue: "Cotton"})
```

```{ojs}
//| echo: false
Plot.plot({
  marginLeft: 55,
  style: {fontSize: "13px"},
  x: {label: "Year", tickFormat: "d"},
  y: {label: "Certificates", grid: true},
  color: {legend: true, domain: sectorDomain, range: sectorRange},
  marks: [
    Plot.barY(transpose(trends_ojs).filter(d => d.sel === sel1),
      {x: "issue_year", y: "n", fill: "sector", tip: true}),
    Plot.ruleY([0])
  ]
})
```

Annual certificates for the selected crop, stacked by applicant category. Ask the cotton
question of your own crop: is it private firms driving the count, or public bodies and
farmers?

## Applicant composition by crop

```{ojs}
//| echo: false
viewof sel2 = pairedCropInput({catValue: "Cereals", cropValue: "Cotton"})
```

```{ojs}
//| echo: false
viewof mode2 = Inputs.radio(["Percentage", "Counts"], {label: "Show", value: "Percentage"})
```

```{ojs}
//| echo: false
rows2 = {
  const all = transpose(comp_ojs).filter(d => d.sel === "All crops").map(d => ({...d, group: "All crops"}));
  const sel = transpose(comp_ojs).filter(d => d.sel === sel2).map(d => ({...d, group: sel2}));
  return all.concat(sel);
}
```

```{ojs}
//| echo: false
Plot.plot({
  marginLeft: 90,
  style: {fontSize: "13px"},
  x: {label: mode2 === "Percentage" ? "Share of certificates" : "Certificates",
      percent: mode2 === "Percentage", grid: true},
  y: {label: null},
  color: {legend: true, domain: sectorDomain, range: sectorRange},
  marks: [
    Plot.barX(rows2, {x: "n", y: "group", fill: "sector",
      offset: mode2 === "Percentage" ? "normalize" : null, tip: true}),
    Plot.ruleX([0])
  ]
})
```

Two thin bars: all crops on top, your selection below. Switch to Percentage to compare
shape rather than scale. Most crops look like the top bar, farmer- or public-heavy. Cotton
is the clearest case of a crop that flips almost entirely to private.

## Firm-level concentration

```{ojs}
//| echo: false
viewof sel3 = pairedCropInput({catValue: "Cereals", cropValue: "Cotton"})
```

```{ojs}
//| echo: false
Plot.plot({
  marginLeft: 220,
  marginRight: 20,
  width: 760,
  style: {fontSize: "13px"},
  x: {label: "Certificates", grid: true},
  y: {label: null},
  marks: [
    Plot.barX(transpose(firm_ojs).filter(d => d.sel === sel3),
      {x: "n", y: "label", sort: {y: "x", reverse: true}, fill: "#377eb8", tip: true}),
    Plot.ruleX([0])
  ]
})
```

The top ten private applicants for the selected crop. Cotton shows Nuziveedu at the top;
other crops show a different set of firms entirely. Use this to check whether
concentration is a cotton-specific pattern or a general one.

## Variety category breakdown

```{ojs}
//| echo: false
viewof sel4 = pairedCropInput({catValue: "Cereals", cropValue: "Cotton"})
```

```{ojs}
//| echo: false
viewof sec4 = Inputs.select(["All"].concat(sectorDomain), {label: "Applicant", value: "Private"})
```

```{ojs}
//| echo: false
rows4 = {
  let r = transpose(var_ojs).filter(d => d.sel === sel4);
  if (sec4 !== "All") r = r.filter(d => d.sector === sec4);
  return r;
}
```

```{ojs}
//| echo: false
Plot.plot({
  marginLeft: 55,
  style: {fontSize: "14px"},
  x: {label: null, tickRotate: -20},
  y: {label: "Certificates", grid: true},
  color: {legend: true, domain: catDomain, range: catRange},
  marks: [
    Plot.barY(rows4, {x: "variety_category", y: "n", fill: "variety_category", tip: true}),
    Plot.ruleY([0])
  ]
})
```

Certificates split across variety categories for the selected crop and applicant. Try
Private together with Cotton: New and Extant (VCK) varieties dominate, while EDVs are
almost invisible.

## Cumulative ownership over time

```{ojs}
//| echo: false
viewof sel5 = pairedCropInput({
  catOptions: ["All crops"].concat(categories), catValue: "All crops",
  cropValue: "Cotton", defaultMode: "category"})
```

```{ojs}
//| echo: false
viewof yr5 = Inputs.range(
  [2009, d3.max(transpose(cum_ojs), d => d.issue_year)],
  {step: 1, label: "Up to year",
   value: d3.max(transpose(cum_ojs), d => d.issue_year)})
```

```{ojs}
//| echo: false
rows5 = transpose(cum_ojs).filter(d => d.sel === sel5 && d.issue_year <= yr5)
```

```{ojs}
//| echo: false
Plot.plot({
  marginLeft: 55,
  style: {fontSize: "13px"},
  x: {label: "Year", tickFormat: "d"},
  y: {label: "Cumulative certificates", grid: true},
  color: {legend: true, domain: sectorDomain, range: sectorRange},
  marks: [
    Plot.line(rows5, {x: "issue_year", y: "cum", stroke: "sector"}),
    Plot.dot(rows5, {x: "issue_year", y: "cum", fill: "sector", r: 3, tip: true}),
    Plot.ruleY([0])
  ]
})
```

Drag the year slider to watch ownership build up rather than just seeing the endpoint.
Default view is all crops; switch to a specific crop to see whether its divergence looks
like cotton's or like the aggregate.

# Cotton in detail

The headline findings established that cotton inverts the usual applicant pattern. This
section supports that finding with more detail on how.

## Ploidy by applicant

```{r}
f9 <- cot %>% filter(!is.na(ploidy), !is.na(sector)) %>% count(sector, ploidy)
ggplot(f9, aes(sector, n, fill = ploidy)) +
  geom_col(position = "dodge") + scale_fill_manual(values = ploidy_cols) +
  geom_text(aes(label = n), position = position_dodge(width = 0.9), vjust = -0.4, size = 3) +
  scale_y_continuous(labels = comma) +
  labs(title = "Cotton certificates by ploidy", x = "Applicant category",
       y = "No. of PVP certificates", fill = NULL)
```

Private cotton certification is mostly tetraploid, the hybrid and Bt lines. Diploid desi
cotton makes up a small share for every type of applicant.

## Variety category by applicant

```{r}
f11 <- cot %>% filter(!is.na(variety_category), !is.na(sector)) %>% count(variety_category, sector)
ggplot(f11, aes(variety_category, n, fill = sector)) + geom_col() +
  scale_fill_manual(values = sector_cols) +
  geom_text(aes(label = n), position = position_stack(vjust = 0.5), size = 3) +
  scale_y_continuous(labels = comma) +
  labs(title = "Cotton certificates by variety category and applicant",
       x = "Variety category", y = "Certificates", fill = NULL) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

## Private cotton, by category over time

```{r}
f12 <- cot %>% filter(applicant_category == "Private", !is.na(variety_category)) %>% count(issue_year, variety_category)
ggplot(f12, aes(issue_year, n, fill = variety_category)) + geom_col() +
  scale_fill_manual(values = cat_cols) +
  geom_text(aes(label = n), position = position_stack(vjust = 0.5), size = 3) +
  scale_x_continuous(breaks = seq(yr_min, max(f12$issue_year), 2)) +
  scale_y_continuous(labels = comma) +
  labs(title = "Private cotton certificates by category",
       x = "Year of certificate issue", y = "Certificates", fill = "Category")
```

---

::: {.callout-tip appearance="minimal"}
Full pipeline (scrape, clean, figures, and this page) is in the
[project repository](https://github.com/ambika-subash/pvp-certification-explorer).
:::
