---
title: "Who owns India's plant varieties?"
subtitle: "Evidence from PPV&FRA certification data"
---

```{r setup}
library(dplyr); library(readr); library(tidyr); library(stringr)
library(ggplot2); library(forcats); library(scales); library(plotly)

d <- read_csv("data/pvp_certificates_clean.csv", show_col_types = FALSE)

yr_min <- 2009L
theme_set(theme_minimal(base_size = 13))
blue <- "#3b6fb6"
cat_cols <- c("Farmer"="#4daf4a","New"="#377eb8","Extant"="#80b1d3",
              "Extant (Notified)"="#ff7f00","Extant (VCK)"="#e41a1c","EDV"="#984ea3")
sector_cols <- c("Farmer"="#4daf4a","Private"="#377eb8","Public"="#e41a1c",
                 "SAU"="#ff7f00","Individual Breeder"="#984ea3")
ploidy_cols <- c("Diploid"="#377eb8","Tetraploid"="#e41a1c")
set_cols    <- c("All crops"="#377eb8","Cotton"="#e41a1c")
sector_levels <- c("Farmer","Private","Public","SAU","Individual Breeder")
cat_levels    <- c("Farmer","New","Extant","Extant (Notified)","Extant (VCK)","EDV")
paper_groups  <- c("Cereals","Fibre Crops","Vegetables","Legumes","Oilseeds","Fruits","Sugar Crops")

d <- d %>% mutate(
  crop_bucket      = ifelse(crop_group %in% paper_groups, crop_group, "Others"),
  sector           = factor(sector, levels = sector_levels),
  variety_category = factor(variety_category, levels = cat_levels))
dd  <- d %>% filter(!is.na(issue_year), issue_year >= yr_min)
cot <- dd %>% filter(is_cotton) %>%
  mutate(ploidy = case_when(
    str_detect(crop, regex("tetraploid", ignore_case = TRUE)) ~ "Tetraploid",
    str_detect(crop, regex("diploid",    ignore_case = TRUE)) ~ "Diploid",
    TRUE ~ NA_character_))

canon_company <- function(x){
  y <- str_to_lower(x); y <- str_replace_all(y, "[.,()]", " ")
  y <- str_replace_all(y, "&", " and "); y <- str_replace_all(y, "\\bseeds\\b", "seed")
  y <- str_replace_all(y, "\\b(private|pvt|limited|ltd|llp|company|co|corporation|corp|incorporated|inc)\\b", " ")
  str_squish(y)
}
alias_company <- function(key) case_when(
  str_detect(key, "maharashtra hybrid seed") | key == "mahyco" ~ "mahyco",
  str_detect(key, "^monsanto") ~ "monsanto", TRUE ~ key)
pretty_names <- c("mahyco" = "Mahyco (Maharashtra Hybrid Seeds)", "monsanto" = "Monsanto")

n_total <- nrow(d)
last_updated <- tryCatch(sub("^Last updated: ", "", readLines("data/last_updated.txt", warn = FALSE)[1]),
                         error = function(e) format(Sys.Date()))
```

::: {.callout-note appearance="minimal"}
**`r format(n_total, big.mark=",")` certificates** analysed, from the
[PPV&FR Authority register](https://plantauthority.gov.in/list-certificates).
Last data refresh: **`r last_updated`**. This page rebuilds every week.
:::

The Protection of Plant Varieties and Farmers' Rights Act (PPVFRA), 2001 set out to
protect breeders' rights while also recognising farmers as innovators. This page reports
what the certification record shows: who secures plant variety protection, for which
crops and categories, and how that has changed over time.

The page is built in three passes. First, **the big picture** across all crops. Then an
**interactive explorer** so you can ask these questions of any crop you care about. Then a
worked case study of **cotton**, India's only widely genetically modified crop, which turns
out to be the clearest example of a pattern that runs through the whole register.

# The big picture

## Certificates issued over time

```{r}
f1 <- dd %>% count(issue_year)
plot_ly(f1, x = ~issue_year, y = ~n, type = "scatter", mode = "lines+markers",
        line = list(color = blue), marker = list(color = blue),
        hovertemplate = "%{x}: %{y} certificates<extra></extra>") %>%
  layout(title = "Total PPVFRA certificates issued, all crops",
         xaxis = list(title = "Year of certificate issue"),
         yaxis = list(title = "No. of certificates"))
```

Certification started slowly, fell during the Covid years, and has since climbed to its
highest levels on record.

## Cotton, on its own

```{r}
f2 <- cot %>% count(issue_year)
plot_ly(f2, x = ~issue_year, y = ~n, type = "scatter", mode = "lines+markers",
        line = list(color = blue), marker = list(color = blue),
        hovertemplate = "%{x}: %{y} certificates<extra></extra>") %>%
  layout(title = "PPVFRA certificates issued, cotton",
         xaxis = list(title = "Year of certificate issue"),
         yaxis = list(title = "No. of certificates"))
```

## What gets certified

```{r}
f3 <- d %>% count(crop_bucket) %>%
  mutate(crop_bucket = fct_relevel(fct_reorder(crop_bucket, n, .desc = TRUE), "Others", after = Inf))
ggplot(f3, aes(crop_bucket, n)) +
  geom_col(fill = blue) + geom_text(aes(label = n), vjust = -0.4, size = 3) +
  labs(title = "Certificates by crop group", x = NULL, y = "Certificates issued") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

Cereals lead by a wide margin. Vegetables are the second-largest group, followed by
fibre crops (which are almost all cotton).

## By variety category

```{r}
f4 <- d %>% filter(!is.na(variety_category)) %>% count(variety_category)
ggplot(f4, aes(variety_category, n, fill = variety_category)) +
  geom_col() + scale_fill_manual(values = cat_cols) + guides(fill = "none") +
  geom_text(aes(label = n), vjust = -0.4, size = 3) +
  labs(title = "Certificates by variety category", x = NULL, y = "Certificates issued")
```

Farmers' varieties are the largest single category overall. The crop by crop breakdown
in the explorer below shows that this total sits very differently across crops.

# Who owns plant varieties?

Certificates are filed by five types of applicants: Farmer (an individual farmer or
farmers' group), Private (for-profit firms), Public (government research bodies), SAU
(State Agricultural Universities), and Individual Breeder (independent plant breeders not
affiliated with any organisation).

## Applicant category, cotton vs all crops

```{r}
f8 <- bind_rows(d %>% count(sector) %>% mutate(set = "All crops"),
                cot %>% count(sector) %>% mutate(set = "Cotton")) %>% filter(!is.na(sector))
ggplot(f8, aes(sector, n, fill = set)) +
  geom_col(position = "dodge") + scale_fill_manual(values = set_cols) +
  geom_text(aes(label = n), position = position_dodge(width = 0.9), vjust = -0.4, size = 3) +
  labs(title = "Applicant category: cotton vs all crops",
       x = "Applicant category", y = "Certificates", fill = NULL)
```

Farmers hold the most certificates across all crops. In cotton the order reverses:
private firms hold almost all of them, and farmers hold close to none.

## Farmers' varieties over time

```{r}
f5 <- dd %>% filter(variety_category == "Farmer") %>% count(issue_year)
ggplot(f5, aes(issue_year, n)) +
  geom_col(fill = cat_cols[["Farmer"]]) + geom_text(aes(label = n), vjust = -0.4, size = 3) +
  scale_x_continuous(breaks = seq(yr_min, max(f5$issue_year), 1)) +
  labs(title = "Certificates issued for farmers' varieties annually",
       x = "Year of certificate issue", y = "Farmers' certificates") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Public applicants, by category

```{r}
f6 <- dd %>% filter(applicant_category == "Public", !is.na(variety_category)) %>% count(issue_year, variety_category)
ggplot(f6, aes(issue_year, n, fill = variety_category)) + geom_col() +
  scale_fill_manual(values = cat_cols) +
  scale_x_continuous(breaks = seq(yr_min, max(f6$issue_year), 2)) +
  labs(title = "Certificates to public applicants, by category",
       x = "Year of certificate issue", y = "Certificates", fill = "Category")
```

## Private applicants, by category

```{r}
f7 <- dd %>% filter(applicant_category == "Private", !is.na(variety_category)) %>% count(issue_year, variety_category)
ggplot(f7, aes(issue_year, n, fill = variety_category)) + geom_col() +
  scale_fill_manual(values = cat_cols) +
  scale_x_continuous(breaks = seq(yr_min, max(f7$issue_year), 2)) +
  labs(title = "Certificates to private applicants, by category",
       x = "Year of certificate issue", y = "Certificates", fill = "Category")
```

## Company concentration across all crops

```{r}
priv_all <- d %>% filter(applicant_category == "Private") %>%
  mutate(key = alias_company(canon_company(applicant)))
labels_all <- priv_all %>% count(key, applicant) %>% group_by(key) %>%
  slice_max(n, n = 1, with_ties = FALSE) %>% ungroup() %>%
  transmute(key, label = coalesce(unname(pretty_names[key]), applicant))
f14 <- priv_all %>% count(key, name = "n") %>% left_join(labels_all, by = "key") %>%
  arrange(desc(n)) %>%
  mutate(label = ifelse(row_number() <= 10, label, "Others")) %>%
  group_by(label) %>% summarise(n = sum(n), .groups = "drop") %>%
  mutate(pct = n / sum(n)) %>% arrange(desc(n))
ggplot(f14, aes("", n, fill = fct_reorder(label, n, .desc = TRUE))) +
  geom_col(width = 1, color = "white") + coord_polar(theta = "y") +
  geom_text(aes(label = ifelse(pct >= 0.03, percent(pct, accuracy = 0.1), "")),
            position = position_stack(vjust = 0.5), size = 3) +
  labs(title = "Major private applicants, all crops", fill = NULL) +
  theme_void(base_size = 12)
```

Market concentration is high across all crops, not just cotton. A small number of
companies dominate private variety certification in India.

# Explore the data yourself

The charts so far describe the register in aggregate. The five tools below let you ask the
same questions of any single crop or crop group. Pick a crop from the dropdown and each
chart re-draws. The cotton story you will read next is just one path through these tools.

```{r}
#| label: prep-ojs
#| include: false

# selection universe = the crop groups plus the ten largest individual crops
top_crops <- d %>% filter(!is.na(crop)) %>% count(crop, sort = TRUE) %>%
  slice_head(n = 10) %>% pull(crop)

# each row is tagged with a "sel" value; rows appear under both their group and
# their crop, so picking "Cereals" or "Rice" both work without double counting
sel_for <- function(df) bind_rows(
  df %>% filter(!is.na(crop_bucket)) %>% mutate(sel = as.character(crop_bucket)),
  df %>% filter(crop %in% top_crops)  %>% mutate(sel = crop))

# OJS serialises characters cleanly; drop factor codes first
chr <- function(df) df %>% mutate(across(where(is.factor), as.character))

trends_df <- chr(sel_for(dd) %>% filter(!is.na(sector)) %>%
  count(sel, issue_year, sector, name = "n"))

comp_df <- chr(sel_for(d) %>% filter(!is.na(sector)) %>%
  count(sel, sector, name = "n"))

firm_key    <- sel_for(d) %>% filter(applicant_category == "Private") %>%
  mutate(firm = alias_company(canon_company(applicant)))
firm_labels <- firm_key %>% count(firm, applicant) %>% group_by(firm) %>%
  slice_max(n, n = 1, with_ties = FALSE) %>% ungroup() %>%
  transmute(firm, label = coalesce(unname(pretty_names[firm]), applicant))
firm_df <- chr(firm_key %>% count(sel, firm, name = "n") %>%
  group_by(sel) %>% slice_max(n, n = 10, with_ties = FALSE) %>% ungroup() %>%
  left_join(firm_labels, by = "firm"))

var_df <- chr(sel_for(d) %>% filter(!is.na(variety_category), !is.na(sector)) %>%
  count(sel, sector, variety_category, name = "n"))

cum_sel <- sel_for(dd) %>% filter(!is.na(sector)) %>% count(sel, issue_year, sector, name = "n")
cum_all <- dd %>% filter(!is.na(sector)) %>% count(issue_year, sector, name = "n") %>%
  mutate(sel = "All crops")
cum_df  <- chr(bind_rows(cum_all, cum_sel) %>% arrange(sel, sector, issue_year) %>%
  group_by(sel, sector) %>% mutate(cum = cumsum(n)) %>% ungroup())

ojs_define(trends_ojs = trends_df)
ojs_define(comp_ojs   = comp_df)
ojs_define(firm_ojs   = firm_df)
ojs_define(var_ojs    = var_df)
ojs_define(cum_ojs    = cum_df)
```

```{ojs}
//| echo: false
sectorDomain = ["Farmer", "Private", "Public", "SAU", "Individual Breeder"]
sectorRange  = ["#4daf4a", "#377eb8", "#e41a1c", "#ff7f00", "#984ea3"]
catDomain = ["Farmer", "New", "Extant", "Extant (Notified)", "Extant (VCK)", "EDV"]
catRange  = ["#4daf4a", "#377eb8", "#80b1d3", "#ff7f00", "#e41a1c", "#984ea3"]
uniqueSel = df => Array.from(new Set(transpose(df).map(d => d.sel))).sort()
```

## Certificate trends by crop

```{ojs}
//| echo: false
viewof sel1 = Inputs.select(uniqueSel(trends_ojs), {label: "Crop / group", value: "Fibre Crops"})
```

```{ojs}
//| echo: false
Plot.plot({
  marginLeft: 55,
  x: {label: "Year", tickFormat: "d"},
  y: {label: "Certificates", grid: true},
  color: {legend: true, domain: sectorDomain, range: sectorRange},
  marks: [
    Plot.barY(transpose(trends_ojs).filter(d => d.sel === sel1),
      {x: "issue_year", y: "n", fill: "sector", tip: true}),
    Plot.ruleY([0])
  ]
})
```

Annual certificates for the selected crop, stacked by applicant category. Ask the cotton
question of your own crop: is it private firms driving the count, or public bodies and
farmers?

## Applicant composition by crop

```{ojs}
//| echo: false
viewof sel2 = Inputs.select(uniqueSel(comp_ojs), {label: "Crop / group", value: "Fibre Crops"})
```

```{ojs}
//| echo: false
viewof mode2 = Inputs.radio(["Counts", "Percentage"], {label: "Show", value: "Percentage"})
```

```{ojs}
//| echo: false
Plot.plot({
  marginLeft: 55,
  x: {label: null},
  y: {label: mode2 === "Percentage" ? "Share of certificates" : "Certificates",
      grid: true, percent: mode2 === "Percentage"},
  color: {legend: true, domain: sectorDomain, range: sectorRange},
  marks: [
    Plot.barY(transpose(comp_ojs).filter(d => d.sel === sel2),
      {x: "sel", y: "n", fill: "sector",
       offset: mode2 === "Percentage" ? "normalize" : null, tip: true}),
    Plot.ruleY([0])
  ]
})
```

Switch to Percentage to see the composition cleanly. Most crops are farmer- or
public-dominated; cotton is the one that flips almost entirely to private.

## Firm-level concentration

```{ojs}
//| echo: false
viewof sel3 = Inputs.select(uniqueSel(firm_ojs), {label: "Crop / group", value: "Fibre Crops"})
```

```{ojs}
//| echo: false
Plot.plot({
  marginLeft: 180,
  x: {label: "Certificates", grid: true},
  y: {label: null},
  marks: [
    Plot.barX(transpose(firm_ojs).filter(d => d.sel === sel3),
      {x: "n", y: "label", sort: {y: "x", reverse: true}, fill: "#377eb8", tip: true}),
    Plot.ruleX([0])
  ]
})
```

The top ten private applicants for the selected crop. Cotton shows Nuziveedu at the top;
maize and rice show different firms. Use this to check whether concentration is unique to
cotton or a general pattern.

## Variety category breakdown

```{ojs}
//| echo: false
viewof sel4 = Inputs.select(uniqueSel(var_ojs), {label: "Crop / group", value: "Fibre Crops"})
```

```{ojs}
//| echo: false
viewof sec4 = Inputs.select(["All"].concat(sectorDomain), {label: "Applicant", value: "Private"})
```

```{ojs}
//| echo: false
rows4 = {
  let r = transpose(var_ojs).filter(d => d.sel === sel4);
  if (sec4 !== "All") r = r.filter(d => d.sector === sec4);
  return r;
}
```

```{ojs}
//| echo: false
Plot.plot({
  marginLeft: 55,
  x: {label: null, tickRotate: -20},
  y: {label: "Certificates", grid: true},
  color: {legend: true, domain: catDomain, range: catRange},
  marks: [
    Plot.barY(rows4, {x: "variety_category", y: "n", fill: "variety_category", tip: true}),
    Plot.ruleY([0])
  ]
})
```

Certificates split across variety categories for the selected crop and applicant. Select
Private together with a cotton option to see how far New and Extant (VCK) varieties
dominate, with EDVs barely present.

## Cumulative ownership over time

```{ojs}
//| echo: false
viewof sel5 = Inputs.select(uniqueSel(cum_ojs), {label: "Crop / group", value: "All crops"})
```

```{ojs}
//| echo: false
viewof yr5 = Inputs.range(
  [2009, d3.max(transpose(cum_ojs), d => d.issue_year)],
  {step: 1, label: "Up to year",
   value: d3.max(transpose(cum_ojs), d => d.issue_year)})
```

```{ojs}
//| echo: false
rows5 = transpose(cum_ojs).filter(d => d.sel === sel5 && d.issue_year <= yr5)
```

```{ojs}
//| echo: false
Plot.plot({
  marginLeft: 55,
  x: {label: "Year", tickFormat: "d"},
  y: {label: "Cumulative certificates", grid: true},
  color: {legend: true, domain: sectorDomain, range: sectorRange},
  marks: [
    Plot.line(rows5, {x: "issue_year", y: "cum", stroke: "sector"}),
    Plot.dot(rows5, {x: "issue_year", y: "cum", fill: "sector", r: 3, tip: true}),
    Plot.ruleY([0])
  ]
})
```

Drag the year slider to watch ownership build up rather than just seeing the endpoint. The
structural divergence between applicant types is a process, not a single snapshot.

# Why cotton is the outlier

Everything in the explorer points the same way: for most crops, farmers and public bodies
hold the bulk of certificates. Cotton is where that structure inverts. This section walks
through why.

## Ploidy by applicant

```{r}
f9 <- cot %>% filter(!is.na(ploidy), !is.na(sector)) %>% count(sector, ploidy)
ggplot(f9, aes(sector, n, fill = ploidy)) +
  geom_col(position = "dodge") + scale_fill_manual(values = ploidy_cols) +
  geom_text(aes(label = n), position = position_dodge(width = 0.9), vjust = -0.4, size = 3) +
  labs(title = "Cotton certificates by ploidy", x = "Applicant category",
       y = "No. of PVP certificates", fill = NULL)
```

Private cotton certification is mostly tetraploid, the hybrid and Bt lines. Diploid desi
cotton makes up a small share for every type of applicant.

## Applicant composition over time

```{r}
f10 <- cot %>% filter(!is.na(sector)) %>% count(issue_year, sector)
plot_ly(f10, x = ~issue_year, y = ~n, color = ~sector, colors = sector_cols,
        type = "scatter", mode = "lines+markers") %>%
  layout(title = "Cotton applicant composition over time",
         xaxis = list(title = "Year of certificate issue"),
         yaxis = list(title = "No. of certificates"))
```

## Variety category by applicant

```{r}
f11 <- cot %>% filter(!is.na(variety_category), !is.na(sector)) %>% count(variety_category, sector)
ggplot(f11, aes(variety_category, n, fill = sector)) + geom_col() +
  scale_fill_manual(values = sector_cols) +
  geom_text(aes(label = n), position = position_stack(vjust = 0.5), size = 3) +
  labs(title = "Cotton certificates by variety category and applicant",
       x = "Variety category", y = "Certificates", fill = NULL) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

## Private cotton, by category over time

```{r}
f12 <- cot %>% filter(applicant_category == "Private", !is.na(variety_category)) %>% count(issue_year, variety_category)
ggplot(f12, aes(issue_year, n, fill = variety_category)) + geom_col() +
  scale_fill_manual(values = cat_cols) +
  geom_text(aes(label = n), position = position_stack(vjust = 0.5), size = 3) +
  scale_x_continuous(breaks = seq(yr_min, max(f12$issue_year), 2)) +
  labs(title = "Private cotton certificates by category",
       x = "Year of certificate issue", y = "Certificates", fill = "Category")
```

## Who are the private cotton firms?

```{r}
priv <- cot %>% filter(applicant_category == "Private") %>%
  mutate(key = alias_company(canon_company(applicant)))
labels <- priv %>% count(key, applicant) %>% group_by(key) %>%
  slice_max(n, n = 1, with_ties = FALSE) %>% ungroup() %>%
  transmute(key, label = coalesce(unname(pretty_names[key]), applicant))
f13 <- priv %>% count(key, name = "n") %>% left_join(labels, by = "key") %>%
  arrange(desc(n)) %>%
  mutate(label = ifelse(row_number() <= 10, label, "Others")) %>%
  group_by(label) %>% summarise(n = sum(n), .groups = "drop") %>%
  mutate(pct = n / sum(n)) %>% arrange(desc(n))
ggplot(f13, aes("", n, fill = fct_reorder(label, n, .desc = TRUE))) +
  geom_col(width = 1, color = "white") + coord_polar(theta = "y") +
  geom_text(aes(label = ifelse(pct >= 0.03, percent(pct, accuracy = 0.1), "")),
            position = position_stack(vjust = 0.5), size = 3) +
  labs(title = "Major private applicants, cotton", fill = NULL) +
  theme_void(base_size = 12)
```

A small number of firms, led by Nuziveedu, hold most of the private cotton certificates.
Varietal control in India's most commercial crop rests with a few companies.

---

::: {.callout-tip appearance="minimal"}
Full pipeline (scrape, clean, figures, and this page) is in the
[project repository](https://github.com/ambika-subash/pvp-certification-explorer).
:::
